---
title: "Covariation"
author: "MPA 634: Data Science for Managers"
date: "9 Feb 2021"
output:
  html_document:
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(plotly)
load(url("http://govfinance.byu.edu/govfinance/classes/dataViz/lectures/illustrations/distributions/distributions.RData"))
```

## Review

### Grammar of Graphics

1.  Data
2.  Aesthetics
3.  Geometric objects
4.  Statistics
5.  Position
6.  Faceting
7.  Coordinate system

In Python, [plotnine](https://pypi.org/project/plotnine/) is a package that implements the grammar of graphics and uses ggplot2 syntax.

[Plotly](%5Bhttps://plotly.com/r/)](<https://plotly.com/r/>)) enhances ggplot.

### Geometric objects for exploration

-   bar
-   column
-   count
-   tile
-   histogram
-   freqpoly
-   density
-   violin
-   boxplot
-   violin
-   point
-   two dimensional density

### Definition of distribution

1.  Range of values
2.  Frequency or probability

### Dimensions of a numerical distribution

1.  Location
2.  Scale
3.  Symmetry
4.  Outliers

## Data frames (tibbles and data sets) to explore

-   mtcars
-   titanic
-   diamonds
-   faithful
-   cows

## mtcars data frame

Objective: Predict mileage using number of cylinders and the weight of the car.

### Summary of data frame

```{r mtcars tibble}
mtcars <- mtcars %>%
  mutate(cyl = factor(cyl)) %>% 
  select(mpg, cyl, wt)
glimpse(mtcars)
summary(mtcars)
```

### Distribution of mileage

#### Summary statistics

```{r distribution of highway mileage}
mtcars %>% 
  summarise(
    Mean = mean(mpg),
    Median = median(mpg),
    Standard_deviation = sd(mpg),
    Interquartile_range = IQR(mpg)
  )
```

#### Violin plot of MPG

```{r violin plot of mpg}
p <- mtcars %>%
  ggplot(aes(y = mpg, x = factor("All Vehicles"))) +
  geom_violin(fill = "lightblue", alpha = 0.5) +
  geom_boxplot(width = 0.10, fill = "grey60") +
  stat_summary(fun = "mean",
    geom = "point",
    shape = 23,
    size = 1.5,
    fill = "red") +
  labs(
    title = "MPG is positively skewed.",
    subtitle = "1973-74 Models",
    y = "Miles Per Gallon",
    caption = "Data Source: R built-in data frame"
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
    ) +
  coord_flip()

ggplotly(p)
```

### Distribution of number of cylinders

```{r distribution of class of car}
mtcars %>%
  ggplot(aes(x = as.factor(cyl), fill = as.factor(cyl))) +
  geom_bar(show.legend = FALSE, alpha = 0.5) +
  labs(
    title = "Few cars in this data set have six cylinders.",
    x = "Number of Cylinders",
    y = "Number of Vehicles"
  ) +
  scale_fill_brewer(palette = "Set2")
```

### Covariation of miles per gallon by number of cylinders using violin plots

```{r covariation of hwy by drive train}
p <- mtcars %>%
  ggplot(aes(x = cyl, y = mpg, fill = cyl)) +
  geom_boxplot(show.legend = FALSE, alpha = 0.25) +
  stat_summary(fun = "mean",
    geom = "point",
    shape = 23,
    size = 1.5,
    fill = "red") +
  labs(title = "Four cyclinder cars get the best mileage.",
       x = "Number of Cylinders",
       y = "Miles Per Gallon") +
  coord_flip()

ggplotly(p)
```

### Covariation between mileage and car weight

```{r mileage and car weight}
p <- mtcars %>%
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point(size = 2) +
  geom_smooth(color = "blue", se = FALSE) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(
    title = "Weight and MPG are inversely related.",
    subtitle = "Nonlinear relationship",
    x = "Weight in 1000 pounds",
    y = "Miles Per Gallon",
    color = "Cylinders",
    caption = "Data Source: Core R"
  )

ggplotly(p)
```

### Covariation of weight and number of cylinders

```{r covariation of weight and number of cylinders}
p <- mtcars %>%
  ggplot(aes(x = cyl, y = wt, fill = cyl)) +
  geom_boxplot(alpha = 0.4) +
  stat_summary(fun = "mean",
    geom = "point",
    shape = 23,
    size = 1.5,
    fill = "red") +
  labs(title = "Four cyclinder cars weight the least.",
       x = "Number of Cylinders",
       y = "Weight in Thousands of Pounds") +
  coord_flip()

ggplotly(p)
```

## [Titanic data set](https://www.kaggle.com/c/titanic)

Objective: Predict whether a Titanic passenger survived base on age and sex.

### Frequency tables

#### Sex

```{r frequency table for sex}
titanic %>% 
  count(sex)
```

#### Survived

```{r frequency table for survived}
titanic %>% 
  count(survived)
```

#### Contingency table for sex and survived using dplyr

```{r contingency}
# Contingency table for diamond color and cut using dplyr and tidyr
titanic %>% 
  count(sex, survived)
```

#### Contingency table for sex and survived using dplyr and tidyr

```{r continency table}
titanic %>% 
  count(sex, survived) %>% 
  spread(sex, n)
```

#### geom_point with position jitter

```{r geom_point}
titanic %>%
  ggplot(aes(x = sex, y = survived, color = survived)) +
  geom_point(position = "jitter",
             shape = 16,
             size = 0.5,
             show.legend = FALSE)
```

#### geom_count

```{r geom_count}
p <- titanic %>%
  ggplot(aes(x = sex, y = survived, color = sex)) +
  geom_count(shape = 15, show.legend = FALSE, alpha = 0.5) +
  scale_size_area(max_size = 45) +
  scale_color_manual(values = c("red", "royalblue"))
ggplotly(p)
```

#### geom_tile

```{r geom_tile}
p <- titanic %>%
  count(sex, survived) %>%
  ggplot(aes(x = sex, y = survived)) +
  geom_tile(aes(fill = n))
ggplotly(p)
```

#### Bar chart

```{r bar chart}
titanic %>% 
  ggplot(aes(y = survived, fill = sex)) +
  geom_bar(position = "fill", alpha = 0.5) +
  labs(
    title = "Female passengers had the highest chance for survival",
    subtitle = "Male passengers likely perished.",
    y = "Survival",
    x = "Percentage",
    caption = "Data Source: Kaggle Titanic Data"
  )
```

### Age of titanic passengers by survival

```{r survival and age}
titanic %>%
  ggplot(aes(x = age, y = survived)) +
  geom_violin(aes(fill = sex), alpha = 0.25, show.legend = FALSE) +
  geom_boxplot(width = 0.30, fill = "grey60", show.legend = FALSE, alpha = 0.5) +
  stat_summary(fun = "mean",
    geom = "point",
    shape = 23,
    size = 1.5,
    fill = "red") +
  facet_grid(rows = vars(sex)) +
  scale_fill_manual(values = c("red", "royalblue")) +
  labs(
    title = "Female survivors were older.",
    subtitle = "Male survivors were younger",
    x = "Age",
    y = "Survived",
    caption = "Source: Kaggle Titanic Data"
  )
```

## diamonds data set

Objective: Predict the price of a diamond based on its weight, cut, color, and clarity.

### Choose a subset of diamonds

```{r smaller diamonds}
set.seed(1234)
dsmall <- diamonds %>%
  sample_n(1000) %>% 
  filter(carat < 3)
```

### Scatterplot of price on carat

```{r scatterplot of carat and price}
cov(dsmall$carat, dsmall$price)
cor(dsmall$carat, dsmall$price)
dsmall %>%
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = 1)
```

### Fixing non linearity and heteroskedasticity with logarithmic transformation

#### Scatterplot with data in original units

```{r original}
dsmall %>%
  ggplot(aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth(color = "blue") + 
  geom_smooth(method = "lm", color = "red")
```

#### Logarithmic transformation

```{r logarithmic transformation}
dsmall %>%
  ggplot(aes(x = log(carat), y = log(price))) +
  geom_point() +
  geom_smooth(color = "blue") + 
  geom_smooth(method = "lm", color = "red")
```

### Two dimensional density

```{r two dimensional density}
dsmall %>%
  ggplot(aes(x = carat, y= price)) +
  geom_density_2d_filled(show.legend = FALSE)
```

### Two dimensional density by cut of diamond

```{r two dimensional density faceted by cut of diamond}
dsmall %>%
  ggplot(aes(x = carat, y= price)) +
  geom_density_2d_filled(show.legend = FALSE) +
  facet_wrap(vars(cut))
```

## Old Faithful data set

Objective: Predict the length of the eruption based on the waiting time.

### Density for eruption length

```{r density for eruption length}
faithful %>%
  ggplot(aes(x = eruptions, y = factor(0))) +
  geom_violin(fill = "lightblue", alpha = 0.5) +
  geom_boxplot(width = 0.10, fill = "grey60") +
  stat_summary(fun = "mean",
    geom = "point",
    shape = 23,
    size = 1.5,
    fill = "red") +
  labs(
    title = "Old Faithful eruption length is bi-modal.",
    x = "Length of Eruption in Minutes",
    caption = "Data Source: Base R data frame"
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
    )
```

### Density for waiting time

```{r density for waiting}
faithful %>%
  ggplot(aes(x = waiting, y = factor(0))) +
  geom_violin(fill = "lightblue", alpha = 0.5) +
  geom_boxplot(width = 0.10, fill = "grey60") +
  stat_summary(fun = "mean",
    geom = "point",
    shape = 23,
    size = 1.5,
    fill = "red") +
  labs(
    title = "Old Faithful eruption length is bi-modal.",
    x = "Length of Waiting Time in Minutes",
    caption = "Data Source: Base R data frame"
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
    )
```

### Scatterplot for waiting and eruptions

```{r scatterplot for waiting and eruption length}
faithful %>%
  ggplot(aes(x = waiting, y = eruptions)) +
  geom_point(position = "jitter") +
  geom_rug(position = "jitter") +
  labs(
    title = "Two clusters exist for waiting and eruption time.",
    subtitle = "Old Faithful waiting and erution times",
    x = "Minutes",
    y = "Minutes",
    caption = "Data source: Base R data frame"
  )
```

### Is there a nonlinear pattern in the scatterplot?

```{r nonlinear pattern}
faithful %>%
ggplot(aes(x = waiting, y = eruptions)) +
  geom_point(position = "jitter") +
  geom_smooth(color = "red", se = FALSE) +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  geom_vline(xintercept = median(faithful$waiting),
             linetype = "dashed",
             color = "gray40") +
  geom_hline(yintercept = median(faithful$eruptions),
             linetype = "dashed",
             color = "gray40") +
  labs(
    title = "Relationship between Old Faithful waiting time and eruption time",
    x = "Waiting Time in Minutes",
    y = "Eruption Duration in Minutes"
  )
```

### Two dimensional density function

```{r two dimensional density for Old Faithful}
faithful %>%
  ggplot(aes(x = waiting, y = eruptions)) +
  geom_density_2d_filled(show.legend = FALSE) +
  labs(
    title = "Relationship between Old Faithful waiting time and eruption time",
    x = "Waiting Time in Minutes",
    y = "Eruption Duration in Minutes"
  )
```

The two dimensional density for faithful uses the viridis color scheme. These colors are chosen to help those who have difficulty distinguishing colors.

## cows data frame

Objective: Predict the butterfat content for a cow given their wage and breed.

### Investigate data frame

```{r cows data frame}
glimpse(cows)
summary(cows)
```

### Covariation of butterfat and breed

```{r covariance of butterfat and breed}
cows %>%
  ggplot(aes(x = butterfat, y = breed %>% fct_reorder(butterfat, median))) +
  geom_violin(fill = "lightblue", alpha = 0.5) +
  geom_boxplot(width = 0.10, fill = "grey60") +
  stat_summary(fun = "mean",
    geom = "point",
    shape = 23,
    size = 1.5,
    fill = "red") +
  labs(
    title = "Jersey and Guernsey cows produce rich milk.",
    subtitle = "Butterfat variability is also high for Jersey and Guernsey cows.",
    x = "Butterfat",
    y = "",
    caption = "Data Source: Wilke Illustrations"
  )
```

### Covariation of butterfat and age

```{r covariance of butterfat and age}
cows %>%
  ggplot(aes(x = butterfat, y = age %>% fct_reorder(butterfat, median))) +
  geom_violin(fill = "lightblue", alpha = 0.5) +
  geom_boxplot(width = 0.10, fill = "grey60") +
  stat_summary(fun = "mean",
    geom = "point",
    shape = 23,
    size = 1.5,
    fill = "red") +
  labs(
    title = "Mature and 2-year old cows have similar butterfat distribution",
    x = "Butterfat",
    y = "",
    caption = "Data Source: Wilke Illustrations"
  )
```

### Cross tabulation of breed and age

```{r cross tab of breed and age}
cows %>% 
  count(breed, age) %>% 
  spread(age, n)
```

### Possible effects of breed and age on butterfat content

```{r covariance of butterfat with breed and age}
cows %>%
  ggplot(aes(x = butterfat, y = breed %>% fct_reorder(butterfat, median))) +
  geom_boxplot(aes(fill = breed), alpha = 0.25, show.legend = FALSE) +
  stat_summary(fun = "mean",
    geom = "point",
    shape = 23,
    size = 1.5,
    fill = "red") +
  labs(
    title = "Cow age interacts with breed to affect butterfat.",
    subtitle = "Two-year olds have more variability in butterfat.",
    x = "Butterfat",
    y = "",
    caption = "Data Source: Wilke Illustrations"
  ) +
  facet_grid(rows = vars(age %>% fct_rev()))
```
