---
title: "Tukey's Five Number Summary"
author: 'MPA 634: Data Science for Managers'
date: "18 Dec 2020"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#### Two parts of a distribution definition

-   Range of Values
-   Frequency

#### Characteristics of a Distribution

-   Location (Central Tendency)
-   Scale (Spread)
-   Symmetry
-   Outliers or tail thickness

Consider the distribution of highway mileage in the mpg data frame which is part of tidyverse. Consider the following values for highway miles per gallon:

```{r highway mile per gallon}
mpg$hwy
```

How do we summarize the distribution of these values?

We can use moment or order based statistics. If we use order based statistics, then we must first order all of the observations:

```{r ordered hwy}
mpg$hwy %>% sort()
```

The quantile function divides that finds the boundary values that leave a certain percentage below. For example, the value `r quantile(mpg$hwy, 0.25)` gives the boundary that has 25% of the observations below that value and 75% of the observations above. This is known as the first quartile.

The second quartile is the middle value. Fifty percent of the observations are below `r quantile(mpg$hwy, 0.5)` which is also known as the median. The third quartile has 75% of the observations below `r quantile(mpg$hwy, 0.75)`.

Measures of Location

Mean

Use the mean function to calculate the average: `r mean(mpg$hwy)`

Median

Use the median function to calculate the median: `r median(mpg$hwy)`

##### Measures of Scale

Standard Deviation

Use the sd function to calculate the standard deviation: `r sd(mpg$hwy)`

Range

Use the range function to calculate the difference between the maximum and the minimum: `r range(mpg$hwy)`

Interquartile range

Interquartile range is the difference between the third and first quartiles. The IQR function calculates the interquartile range: `r IQR(mpg$hwy)`

#### Tukey's Five Number Summary of a numerical variable includes:

1.  Minimum
2.  First Quartile
3.  Median (Second Quartile)
4.  Third Quartile
5.  Maximum

Summary command in R gives Tukey's five number summary and the mean: 

```{r summary of hwy}
summary(mpg$hwy)
```

John Tukey invented boxplots to depict the five number summary graphically.



