---
title: "Final Exam Winter 2021"
author: "Data Science for Managers"
date: "Due by midnight: 21 April 2020"
output:
  html_document:
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(nycflights13)
library(leaflet)
```

Exam Parameters

-   Open book but closed other persons
-   No time limit

Submit RMD and HTML files that contain code for the following five problems. You shouldn't use any software other than R to complete your calculations. Remember that if you have an error in your code chunk, you can still get your file to knit if you put error = TRUE in the code chunk header.

### Problem 1: Violin plot of two numerical variables

Create a real violin plot for the hwy and cty variables from the mpg tibble.

-   Convert the wide mpg tibble into one that is long
-   Create a violin and boxplot combination with no legend
-   Include the mean as part of your violin plot
-   Use the Pastel1 brewer palette
-   Carefully label the graph so that yours matches the violin plot shown.

```{r Problem 1}
# put your code here
```

### Problem 2: Top 5 destinations from JFK

-   Calculate the number of flights leaving from JFK for each destination airport.
-   Sort the destinations in descending order and choose the top 5 destinations.
-   Save the tibble so that it can be used in the next problem.
-   Print the top five airports.

```{r Problem 2}
# put your code here
```

### Problem 3: Bar chart with filtering

-   Use the tibble from Problem 2 and a filtering join to select only those rows from the flights tibble that fly to the top five destinations. If you haven't been able to complete Problem 2 successfully, then email me and I will send you the tibble you need to complete this problem.
-   Choose only those flights that leave from JFK.
-   Create a horizontal Pareto chart with the largest frequency (LAX) at the top of the graph.
-   Use the Pastel2 brewer palette.
-   Optionally, if you are looking for an additional challenge, you could also use a mutating join to add the airport names to your graph rather than the codes (LAX is Los Angeles).

```{r Problem 3}
# put your code here
```

### Problem 4: Largest deviation

Write a function that identifies the size of the most deviant observations in a numerical vector as defined by the largest absolute value of the z-score.

-   Calculate the deviation from the mean.
-   Calculate the z-score by dividing each deviation by the standard deviation.
-   Calculate the absolute value of each z-score.
-   Find the largest absolute z-score and round the result to two decimal places.
-   Test your function on the hwy column from the mpg tibble.
-   Extra Credit: Show that you can accomplish the same thing with one line of code in the body by using the **scale()** function.

```{r Problem 4}
# put your code here
```

The z-score for the most deviant observation in the hwy column in the mpg tibble is: `r most_deviant(mpg$hwy)`

```{r Problem 4 one line body}
# put your code here
```

The z-score for the most deviant observation in the hwy column in the mpg tibble that is found by using the scale function is: `r most_deviant_efficient(mpg$hwy)`

### Problem 5: Seattle airbnb

-   Import the seattle airbnb data from Kaggle
-   Take a sample of 500 properties from the tibble
-   Use the map set Esri.NatGeoWorldMap
-   Add blue circle markers with radius 4 and opacity of 0.2

In order to knit your exam file, put the file that has the Seattle data in the same folder as your coding problems rmd file.

```{r seattle map}
# put your code here
```
