---
title: "R4DS"
author: "Ray Nelson"
date: "6/4/2018"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
theme_set(theme_grey())
```

## 3.2 First steps

### 3.2.1 The mpg data frame

```{r mpg data frame}
mpg
```  
Typing in the name of an object outputs the object to the console.

### 3.2.2 Creating a ggplot

```{r ggplot of highway vs displacement}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```  

### 3.2.4 Exercises

#### 1. Run ggplot(data = mpg). What do you see?  
```{r 3.2 problem 1}
ggplot(data = mpg)
```
This creates a graph but doesn't put any information on the graph using the layers that are inherent in geom constructs.  

#### 2. What are the dimensions of mpg?  
```{r dimensions of mpg}
mpg %>% dim()
mpg %>% ncol()
mpg %>% nrow()
```

#### 2. What does the drv variable describe?
```{r drv variable}
? mpg
```  

#### 4. Make a scatterplot of hwy versus cyl
```{r scatterplot of numerical variables, continuous versus discrete}
ggplot(data = mpg) +
  geom_point(aes(x = cyl, y = hwy))
```  

#### 5. What happens if you make a scatterplot of class vs drv?  

```{r scatterplot of two categorical variables}
ggplot(data = mpg) +
  geom_point(aes(x = drv, y = class))
```  

## 3.3 Aesthetic mappings

```{r aesthetics introductory mapping}
ggplot(data = mpg %>% filter(class != "2seater")) +
  geom_point(aes(x = displ, y = hwy)) +
  geom_point(data = mpg %>% filter(class == "2seater"),
             aes(x = displ, y = hwy), color = "red") +
  geom_smooth(aes(x = displ, y = hwy), method = "loess")
```  
Possible aesthetics include size, shape, color,fill, opacity, or linetype.  

```{r class of vehicle as an aesthetic}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  labs(
    title = "Scatterplot of Mileage and Engine Displacement",
    x = "Displacement",
    y = "Highway MPG",
    color = "Class of Vehicle"
  )
```  

```{r class as a size aesthetic}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```  

```{r class as transparency and shape aesthetics}
left <- ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class)) +
  labs(
    title = "Transparency aesthetic",
    x = "Displacement",
    y = "MPG"
  )
right <- ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class)) +
  labs(
    title = "Shape aesthetic",
    x = "Displacement",
    y = "MPG"
  )
plot_grid(left, right, nrow = 2, align = "v")
```

```{r arbitrary color for points}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

#### 3.3.1 Exercises  

##### 1. What's gone wrong with this code?
```{r why points are blue}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```  

##### 2. Which variables in mpg are categorical? Which variables are continuous?

```{r types of variables}
? mpg
mpg %>% summary()
```  

##### 3.  Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical versus continuous variables?  

```{r categorical and continuous aesthetics}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = hwy))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = hwy))

#ggplot(data = mpg) +
  #geom_point(mapping = aes(x = displ, y = hwy, shape = hwy))
```  
Shape can't be mapped to a continuous variable. The mapping for color and size is possible.  

##### 4. What happens if you map the same variable to multiple aesthetics?

```{r same variable to multiple aesthetics}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = fl, shape = fl))
```  

##### 5. What does the stroke aesthetic do? What shapes does it work with?  

```{r stroke aesthetic}
? geom_point
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class), stroke = 2, shape = 21)
```  

##### 6. What happens when you map an aesthetic to something other than a variable name.

```{r mapping to something other than a variable name}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```

#### 3.5 Facets

```{r facet example}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```  

```{r facet example with two variables}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(drv ~ cyl)
``` 

##### 3.5.1 Exercises

##### 1. What happens if you facet on a continuous variable?  

```{r faceting on a continuous variable}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(displ ~ drv)
```  
A plot is create for each value of the continuous variable. This isn't very helpful.  

##### 2. What do the empty cells in the plot with facet_grid(drv ~ cyl) mean?  

```{r empty cells}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```  
The empty cells in the grid indicate the the combinations of drive and cylinders that don't contain any observations.

##### 3. What plots do the the follow code create? What does . do?  

```{r effect of the . notation}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```  
The . notation gives a grid with only one category. The variable to the left of the ~ gives the rows of the grid, just like matrix notation. The variable to the left of the ~ gives the columns. The . refers to all possible rows that satisfy the dimension in the column.  

##### 4. What are the advantages and disadvantages of the first faceted plot in this section?  

```{r advantages of faceting versus color}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```  
Facets allow analysts to concentrate on the patterns for each category without the distractions of all the other data.

##### 5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol argument?  

With facet_wrap, the number of columns or rows that are created before starting to wrap are given the nrow and ncol parameters. With the facet_grid, the number of columns and rows is determined by the number of categories for the variable(s) of interest.

##### 6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

The larger number of categories should be rows because we have more flexibility to create a large number of rows that columns.  

#### Geometric objects

```{r different geometric objects, message = FALSE}
# left
left <- ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# right
right <- ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))

plot_grid(left, right, align = "h")
```
 
```{r linetype aesthetic, message = FALSE}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r overlay of point and smooth, message = FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv, linetype = drv)) + 
  geom_smooth() +
  geom_point()
```

```{r group aesthetic}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
              
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```






