---
title: 'Homework #2 Key'
author: "Aesthetic Mappings and Facets"
date: "Due: 19 Jan 2021"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
```

## Concepts

Explain how aesthetics, scaling, and mapping are related to each other within the context of the grammar of graphcs and ggplot.

Aesthetics are visual properties in a graph. They include x-axis, y-axis, color, fill, linetype, transparency, shape, and stroke. Scaling and mapping are referring to the same concept. Mapping and scaling assign levels to elements in a graph. These levels can be assigned based on variables in a data frame. They can also be assigned values that apply globally to all observations.

## Interpretation of code

```{r code chunk}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point(show.legend = FALSE) +
  geom_smooth(mapping = aes(linetype = drv), method = "lm", se = FALSE, show.legend = FALSE) +
  facet_wrap(vars(drv))
```

Line 1: calls the ggplot function. It gets the data from the data frame (tibble) mpg and assigns displacement to the x-axis, highway miles per gallon to the y axis, and associates color with the vehicle's drive train.

Line 2: draws a scatterplot. It suppresses the legend that would result because of the color = drv statement.

Lines 3: creates a simple regression line whose color and linetype are determined by the type of drive train. The method = "lm" argument chooses the smoothing method to be simple regression. The se = FALSE clause removes the confidence interval from the graph. The show.legend = FALSE once again prevents a legend from being drawn.

Line 4: creates a separate scatterplot for each class of vehicle

## Write your own code

Draw a scatterplot of linear regressions of city mileage on displacement

1.  Put displ on the x-axis
2.  Put cty on the y-axis
3.  Color the plot symbols by class
4.  Draw a simple regression line for each car class without a confidence interval
5.  Use facet_wrap to create a graph for each class
6.  Suppress the legend

```{r facets}
ggplot(data = mpg, mapping = aes(x = displ, y = cty, color = class)) +
  geom_point(show.legend = FALSE) +
  geom_smooth(method = "lm", se = FALSE, show.legend = FALSE) +
  facet_wrap(vars(class), ncol = 2)
```
