---
title: 'Homework #4'
author: "Ray Nelson"
date: "1/16/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Concepts  

Explain the sequence whereby a word file is created from a R Markdown document. What is the purpose of the YAML header?

## Interpretation of code  

Interpret the following R code:  

```{r code chunk 1}
ggplot(data = faithful, mapping = aes(x = eruptions)) +
  geom_histogram(aes(y = stat(density)), fill = "lightgreen") +
  geom_density(fill = "blue", alpha = 0.3) +
  geom_rug(color = "red")
``` 

```{r code chunk 2}
ggplot(data = mpg, mapping = aes(x = class, y = hwy, fill = class)) +
  geom_violin(show.legend = FALSE) +
  geom_boxplot(show.legend = FALSE, width = 0.05, fill = "grey")
```

## Compare the distribution of city mileage from the mpg dataframe for each of the three different types of drive trains using:

1.  Facets with a density superimposed on top of a histogram
1.  Violin plots

```{r facets}

```

```{r violin plots}

```

