---
title: 'Homework #5'
author: "Filtering and Arranging"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

## Concepts  

1.  Explain the differences among "=", "==", and "<-". Give an example of how you would use each of these symbols.  
1.  Compare and contrast the logical operators "x & y" and "xor(x, y)".

## Interpretation of code  

Interpret the following R code:  

```{r code chunk}
delta_spring_fall <- filter(flights, month == 5 | month == 9, carrier == "DL")
ggplot(data = filter(delta_spring_fall, dep_delay < 90),
       mapping = aes(x = as.factor(month),
                     y = dep_delay,
                     fill = as.factor(month))) +
  geom_violin(show.legend = FALSE) +
  geom_boxplot(width = 0.05, fill = "grey", show.legend = FALSE) +
  coord_flip() +
  facet_grid(rows = vars(origin))
```

## Coding Problem 1: Use violin plots and facets to explore the distribution of departure delays for the American, Delta, and United (AA, DL, and UA) for flights going to Atlanta, Chicago, and Los Angeles (ATL, ORD, and LAX).

```{r airlines and destinations}

```

## Coding Problem 2: Air time and distance travelled are undoubtedly positively related. Investigate this relationship for all April flights from JFK.
*  Use jittering and transparency to help with the overplotting.
*  What might explain the clusters of points in the scatterplot?
*  Is the relationship linear?

```{r air time and distance travelled}

```

