---
title: 'Homework #6'
author: "Select and Mutate"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

## Concepts  

1.  Compare and contrast filter and select.  
1.  Explain how ":" can be used when selecting variables.

## Interpretation of code  

Interpret the following R code:  

```{r code chunk interpretation, warning = FALSE}
converted_times <- mutate(
  flights,
  hours = dep_time %/% 100,
  minutes = dep_time %% 100,
  minutes_after_midnight = 60 * hours + minutes
)
ggplot(
  data = converted_times,
  mapping = aes(x = origin,
                y = minutes_after_midnight,
                fill = origin)
) +
  geom_violin(show.legend = FALSE) +
  geom_boxplot(width = 0.1,
               fill = "grey",
               show.legend = FALSE) +
  coord_flip()
```

## Coding Problem 1:

1.  Select only those flights going from JFK to Chicago.  
1.  Create a data frame that includes the following variables: dep_delay and arr_delay.    
1.  Rename the variables dep_delay to departure and arr_delay to arrival.  
1.  Draw a scatterplot of arrival on departure with colored loess and linear regression smoothers without confidence intervals.  


```{r JFK to Chicago}

```  

## Coding Problem 2: Use mutate and the ranking function to find the 5 most delayed flights departing from JFK. Print out only the rank of the delay, year, month, day, origin, dest, carrier, sched_dep_time, dep_delay, and delay with the most delayed at the top of the table.  

```{r 5 most delayed flights}

```  

