---
title: 'Homework #7'
author: 'Summarization by Group'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

## Concepts  

Carefully explain the following:
1.  The two elements that define the distribution of a random variable
1.  How the mean measures the location of a distribution
1.  How the standard deviation measures the scale or spread of a distribution

## Interpretation of code  

Interpret the following R code:  

```{r code chunk interpretation, warning = FALSE, message = FALSE}
flights %>%
  group_by(day, month) %>%
  summarize(cancelled = mean(is.na(dep_time)) * 100,
            delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(mapping = aes(x = delay, y = cancelled)) +
  geom_point() +
  geom_smooth(color = "red") +
  labs(
    title = "Large average delays associate with percentage of cancelled flights",
    x = "Average Delay in Minutes",
    y = "Percentage of cancelled flights"
  )
```

## Write code that accomplishes the following:

1.  Create a new data frame that excludes cancelled flights
1.  Rank carriers from best to worst by their average arrival delays
1.  Rank destination airports from worst to best by their average arrival delays
1.  Rank NYC airports from best to worst by their average departure delays
1.  Rank carriers from best to worst by their on time percentage

Is it possible to use the above results to rank carriers? As a hint, think about how the following code might help you consider this question.  

```{r number of flights}
flights %>% 
  group_by(carrier, origin) %>%
  count(dest)
```  