---
title: 'Homework #8'
author: 'Univariate Distributions'
date: "Due: 9 Feb 2021"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
```

## Concepts

1.  Carefully explain how visualizing, transforming, and modeling interact in exploratory data analysis.

2.  Identify geometric objects that appropriately depict the distributions of

    a.  Categorical variables
    b.  Numerical variables

## Code Interpretation

```{r code chunk interpretation}
mpg %>% 
  filter(year == 1999) %>% 
  ggplot(aes(y = factor("City"), x = cty)) +
  geom_violin(fill = "lightblue", alpha = 0.4) +
  geom_vline(xintercept = median(mpg$cty), color = "red", linetype = "dashed") +
  geom_vline(xintercept = quantile(mpg$cty, 0.25), color = "blue", linetype = "dashed") +
  geom_vline(xintercept = quantile(mpg$cty, 0.75), color = "blue", linetype = "dashed") +
  geom_boxplot(fill = "lightgreen", width = 0.25, alpha = 0.4) +
  stat_summary(fun = "mean",
     geom = "point",
     shape = 23,
     size = 1.5,
     fill = "red") +
  labs(
    title = "City mileage may not be normally distributed.",
    subtitle = "Model Year 1999",
    y = "",
    x = "Miles Per Gallon",
    caption = "Data source: Tidyverse mpg"
  )
```

## Write code

The faithful dataset includes 272 observations that report the length of time that the Old Faithful geyser in Yellowstone National Park erupts and the time interval between the eruptions. Use exploratory data analysis to investigate the distributions of each of the two variables, eruptions and waiting. Your code should generate summary statistics and graphs that give you insights into:

1.  Location
2.  Scale
3.  Symmetry
4.  Outliers
