---
title: 'Homework #9'
author: 'Covariation'
date: "Due: 11 Feb 2021"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
library(plotly)
```

## Concepts

##### 1. Identify at least five geom's that help you explore the covariation between two variables.

##### 2. Explain how you can use violin plots to determine whether a distribution is symmetrical.

##### 3. Explain how you can use violin plots to determine whether a distribution has fat tails relative to the normal.

## Interpretation of code

##### Problem 1

```{r Problem 1 code chunk interpretation}
faithful %>%
  ggplot(aes(x = eruptions, y = factor(0))) +
  geom_violin(fill = "lightblue", alpha = 0.5) +
  geom_boxplot(width = 0.10, fill = "grey60") +
  stat_summary(fun = "mean",
    geom = "point",
    shape = 23,
    size = 1.5,
    fill = "red") +
  labs(
    title = "Old Faithful eruption length is bi-modal.",
    x = "Length of Eruption in Minutes",
    caption = "Data Source: Base R data frame"
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
    )
```

##### Problem 2

```{r Problem 2 code chunk interpretation}
p <- mtcars %>%
  ggplot(aes(x = wt, y = mpg, color = as.factor(cyl))) +
  geom_point(size = 2) +
  geom_smooth(color = "blue", se = FALSE) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(
    title = "Weight and MPG are inversely related.",
    subtitle = "Nonlinear relationship",
    x = "Weight in 1000 pounds",
    y = "Miles Per Gallon",
    color = "Cylinders",
    caption = "Data Source: Core R"
  )

ggplotly(p)
```

## Write code

Use the mpg data frame and ggplot to investigate the covariation between

##### 1. Highway (hwy) and city (cty) mileage

```{r relationship between highway and city mileage}
# put your code here
```

##### 2. Highway mileage and drive train

```{r relationship betwen highway and drive train}
# put your code here
```
