---
title: 'Homework #9'
author: 'Covariation'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Concepts  

1.  Under what conditions is it appropriate to use covariance and correlation as measures of the strength of relationships between variables.

    
## Interpretation of code  

Interpret the following R code:  

```{r code chunk interpretation I}
mpg %>%
  ggplot(aes(x = reorder(class, cty, FUN = IQR), y = hwy)) +
  geom_violin(fill = "lightgreen") +
  geom_boxplot(width = 0.1, fill = "lightblue") +
  labs(
    title = "Violin Plot of City Miles Per Gallon",
    x = "Class of Vehicle",
    y = "Miles Per Gallon") +
  coord_flip()
```

```{r code chunk interpretation II}
diamonds %>%
  filter(carat < 3) %>% 
  ggplot(aes(x = carat, y = price)) +
  geom_boxplot(aes(group = cut_width(carat, 0.1))) +
  coord_flip()
```


## Write code  

Using the mpg data set, use covariance, correlation, scatterplots, and heat maps to investigate the relationship between highway and city mileage. Use geom_smooth to determine whether the relationship is linear.

```{r relationship between highway and city mileage}

```



