---
title: 'Homework #9'
author: 'Covariation'
date: "Due: 7 Oct 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Concepts  

#### 1.  Explain how Tukey's five number summary is related to the boxplot.

#### 2.  Identify at least five geom's that help you explore the covariation between two variables.
  
#### 3. Explain how you can use violin plots to determine whether a distribution is symmetrical.

#### 4. Explain how you can use violin plots to determine whether a distribution has fat tail relative to the normal.
    
## Interpretation of code  

```{r code chunk interpretation}
mpg %>%
  ggplot(aes(x = reorder(class, cty, FUN = IQR), y = cty)) +
  geom_violin(fill = "lightgreen") +
  geom_boxplot(width = 0.1, fill = "lightblue") +
  labs(
    title = "Violin Plot of City Miles Per Gallon",
    x = "Class of Vehicle",
    y = "Miles Per Gallon") +
  coord_flip()
```  

## Write code  

Use the mpg data frame and ggplot to investigate the covariation between

#### 1. Highway (hwy) and city (cty) mileage

```{r relationship between highway and city mileage}

```

#### 2. Highway mileage and drive train

```{r relationship betwen highway and drive train}

```




