---
title: 'Homework #10'
author: 'Data Wrangling'
date: "Due: 25 Feb 2021"
output:
  html_document:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
```

## Find a dataset of interest from Kaggle, import the data into R, and then explore the following:

1.  The distribution of a categorical variable.
1.  The distribution of a numerical variable.
1.  Covariation between a categorical variable and a numerical variable
1.  Covariation between two numerical variables

## Import data from Kaggle

```{r import data into a tibble}

```


## Distribution of categorical variable

```{r categorical distribution}

```

## Distribution of a numerical variable  

```{r numerical distribution }

```


## Covariation between a categorical variable and a numerical variable

```{r covariation categorical and numerical}

```


## Covariation between two numerical variables  

```{r covariation two numerical variables}

```

