---
title: 'Homework #11'
author: 'Tidy Tibbles'
date: "Due: 2 March 2021"
output:
  html_document:
    code_folding: show
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(readxl)
```

## Concepts

1.  Explain why pivot_longer and pivot_wider are opposite operations.
2.  Explain why separate and unite are opposite operations.

## Code Interpretation

Interpret the following R code chunk:

```{r code chunk interpretation 1}
who %>% 
  pivot_longer(
    cols = new_sp_m014:newrel_f65,
    names_to = "key",
    values_to = "cases",
    values_drop_na = TRUE
  ) %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel")) %>%
  separate(col = key,
           into = c("new", "type", "sexage")) %>%
  select(-new, -iso2, -iso3) %>% 
  separate(
    col = sexage,
    into = c("sex", "age"),
    sep = 1
  )
```

## Write code

Complete the following steps:

1.  Import the excel file interest_rates into R.
2.  Use pivot_longer to create a long tibble.
3.  Draw a time series plot with both interest rates on the same graph. Put date on the x-axis. Put mortgage and treasury on the y-axis.
4.  Use a boxplot to compare the distributions of the 30 year mortgage and 30 year Treasury bonds.

```{r interest rates}
interest_rates <- read_excel("Interest rates.xlsx")

```
