---
title: 'Homework #13'
author: 'Factors'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
```

## Concepts  

1.  How are factors different from character variables. When would you prefer factors?

2.  Explain how you can change

*  Factor order
*  Factor levels
    
## Interpretation of code  

Compare and interpret the following two R code chunks: 

### Code chunk #1  

```{r code chunk interpretation 1, warning = FALSE, message = FALSE}
gss_cat %>%
  mutate(marital = marital %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(x = marital, fill = marital)) +
  geom_bar(show.legend = FALSE) +
  labs(
    title = "Bar chart with factors ordered by frequency",
    subtitle = "This creates a Pareto Chart.",
    x = "",
    y = "Number of Respondents"
  )
```  

### Code chunk #2

```{r code chunk interpretation 2, message = FALSE}
gss_cat %>%
  mutate(partyid = fct_collapse(
    partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  )) %>%
  ggplot(aes(x = partyid, fill = partyid)) +
  geom_bar(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Independents are the largest group of respondents",
    x = "",
    y = "Number of Respondents"
  )
```

## Write code

Create a Pareto bar chart for race with the largest category on left side of the graph. Remember that fct_infreq will help you reorder the factor race.

```{r Pareto Chart}

```  

### Problem 2

Create a boxplot for age by marital status. Sort the boxes by the median of age. The following steps will help you construct this graph:

1.  Use the gss_cat tibble
2.  Select the variables marital and age
3.  Use piping and na.omit to get rid of missing observations
4.  Draw a boxplot with the boxes colored by marital status
5.  Reorder the boxes using fct_reorder (the default function is median)
6.  Suppress the legend
7.  Create appropriate titles and axis labels

```{r boxplot of age by marital status, warning = FALSE}

```  
