---
title: 'Homework #15'
author: 'Functions'
date: "Due: 9 March 2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
library(lubridate)
library(nycflights13)
```

## Concepts  

1.  Give two original examples of functions where you have the function name as a verb and the arguments as nouns.  

2.  Explain why style is important when you are writing functions.

3.  What is meant by snake_case and camelCase when you are writing functions? 
    
## Interpretation of code  

Interpret the following R code chunk:  

```{r code chunk interpretation}
graph_dest_hours <- function(dest_airport){
flights %>% 
  filter(dest == dest_airport) %>%
  ggplot(aes(hour, fill = origin)) +
  geom_bar() +
  scale_x_continuous(breaks = seq(0, 24), labels = seq(0, 24)) +
  labs(
    title = paste("New York City Flights to", dest_airport),
    x = "Hour of the day",
    y = "Number of flights",
    fill = "New York Airport"
  )  
}

graph_dest_hours("LAX")
```  

## Write Code

### Problem 1  

Write code that creates a function that takes as an argument the origin airport in the flights tibble that will to accomplish the following

1.  Remove observations with missing values for dep_time
1.  Select all departing flights for origin airport
1.  Calculate the hour from dep_time using integer division
1.  Draw a bar chart for the new hour variable
1.  Optionally, you can creating labeling for the graph.

Run your function for the JFK airport.

```{r homework image, echo = FALSE}
homework_graph <- "http://govfinance.byu.edu/govfinance/classes/dataScience/homework/graphs/HW15 Graph.png"
```

![](`r homework_graph`)

```{r departure function}

```

### Problem 2

Write a function that calculates the number of days from today until the last day of class. Your function should take as an argument the last day of class. 

```{r days until the end of the semester}

```
