---
title: 'Homework #16'
author: 'Vectors, Lists, and Tibbles'
date: "Due: 25 March 2021"
output:
   html_document:
     code_folding: show
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
```

## Concepts

1.  Compare and contrast atomic vectors and lists.

2.  Compare and contrast implicit and explicit coercion.

## Interpretation of code

### Code Chunk \#1

```{r atomic vectors into a list and tibble}
age <- c(24L, 30L)
income <- c(24, 50)
first_name <- rep("BYU", 2)
last_name <- paste0("Student", 1:2)
list(first_name, last_name, age, income)
tibble(first_name, last_name, age, income)
```

### Code Chunk \#2

```{r character integers and numbers}
(character_numbers <- rep(c("1", "2", "3", "4"), 3))
character_numbers %>% as.integer()
character_numbers %>% as.integer() %>% typeof()
(integers <- rep(1:4, 3))
typeof(integers)
(real_numbers <- as.numeric(integers))
typeof(real_numbers)
```

## Write Code

For your family,

-   Create a vector that includes the first name of each member.
-   Create a vector that includes their birthdays.
-   Create a vector of the addresses where your family has lived.
-   Create a list of the first names, birthdays, and addresses.
-   Create a tibble of the first names and birthdays.
