---
title: 'Homework #16 Key'
author: 'Vectors, Lists, and Tibbles'
date: "Due: 25 March 2021"
output:
   html_document:
     code_folding: show
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
```

## Concepts

1.  Compare and contrast atomic vectors and lists.

    Atomic vectors are homogeneous. This means that they must contain the same type of data. The major types of data are logical, integer, double, and character. All atomic vectors have a length.

    Lists are heterogeneous repositories of information. This means that lists can contain more than one type of object. The slots in a list can contain atomic vectors of any length and other lists. Lists are sometimes called recursive vectors because they can contain other lists.

    Homogeneity for atomic vectors and heterogeneity for lists are the major difference between the two.

2.  Compare and contrast implicit and explicit coercion.

    Coercion is when we change the type of a vector. For example, we may want to change the integer vector (1, 2, 3, 4) into a character vector ("1", "2", "3", "4").

    We can change the type of the vector explicitly by using a function such as as.character in the previous example.

    Sometimes vectors are changed as part of calculation so the conversion is built into the functions that we are calling. This is implicit coercion.

## Interpretation of code

### Code Chunk \#1

```{r atomic vectors into a list and tibble}
age <- c(24L, 30L)
income <- c(24, 50)
first_name <- rep("BYU", 2)
last_name <- paste0("Student", 1:2)
list(first_name, last_name, age, income)
tibble(first_name, last_name, age, income)
```

Line 1: creates an integer atomic vector called age of length 2. It puts 24 in the first slot and 30 in the second slot.

Line 2: creates a double atomic vector called income of length 2. It puts 24 in the first slot and 50 in the second slot.

Line 3: creates a character atomic vector called first_name of length 2. Both the first and second slots contain the value BYU.

Line 4: creates a character atomic vector called last_name of length 2. The paste0 function combines the values "Student" with the numbers 1 and 2. The process recycles the "Student" value so that there are two instances of "Student" to go with the numbers 1 and 2. The paste0 function doesn't put any spaces between the values being pasted together.

Line 5: creates a list of length 4. The vectors first_name, last_name, age, and income go into the first through fourth slots in the vector.

Line 6: creates a tibble that is composed of four atomic vectors of the same length. If the vectors didn't have the same length, then we wouldn't be able to create the tibble. A tibble is a list of atomic vectors that have the same length.

### Code Chunk \#2

```{r character integers and numbers}
(character_numbers <- rep(c("1", "2", "3", "4"), 3))
character_numbers %>% as.integer()
character_numbers %>% as.integer() %>% typeof()
(integers <- rep(1:4, 3))
typeof(integers)
(real_numbers <- as.numeric(integers))
typeof(real_numbers)
```

Line 1: creates a character atomic vector of length 12 which contains the strings "1", "2", "3", "4" repeated three times. The () cause the result to be printed to the console.

Line 2: explicitly coerces the character_numbers from character to integer

Line 3: determine the type of the result of the coercion in line 2. It returns the value of integer.

Line 4: creates an integer atomic vector of length 12 by repeating the sequence of number 1, 2, 3, 4 three times.

Line 5: reports that the type of the vector integers is integer.

Line 6: explicitly coerces the integer atomic vector into into a double

Line 7: reports that the result of the coercion is double

## Write Code

For your family,

-   Create a vector that includes the first name of each member.
-   Create a vector that includes their birthdays.
-   Create a vector of the addresses where your family has lived.
-   Create a list of the first names, birthdays, and addresses.
-   Create a tibble of the first names and birthdays.

```{r family list and tibble}
nelsons <- c("Ray", "Pam", "Jen", "Bec", "Ang", "Jon")
nelson_birthdays <- c("14-Apr", "20-Apr", "13-Mar", "26-Jan",
                     "5-Mar", "18-Jun")
nelson_addresses <- c("115 Huerta Place, Davis, California",
                      "513 East 4380 North, Provo, Utah",
                    "836 East 2680 North, Provo, Utah")
nelson_list <- list(nelsons, nelson_birthdays, nelson_addresses)
nelson_tibble <- tibble(nelsons, nelson_birthdays)
```
