---
title: 'Homework #17'
author: "Lists and methods"
date: "Due: 30 March 2021"
output: 
  html_document:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
```

## Concepts

1.  Explain why a tibble is a list but not all lists are tibbles.
2.  Explain how you use [], [[]], and \$ when subsetting a list as described in section 20.5.2.
3.  Explain how class and method are related to each other. Use the summary method as an example to explain your answer.

## Interpretation of code

Interpret the following R code chunk:

```{r extracting information from lists}
data_science <- list(transform = "Transform", model = "Model", visual = "Visual")
our_list <- list(one = TRUE, two = 3L, three = 500, four = "EDA", five = data_science)
our_list$five
our_list[1]
our_list[[1]]
our_list[["one"]]
our_list[1:2]
our_list$five$model
```

## Write Code

Complete the following steps

-   Create a new object for a linear regression of cty on displ in the mpg tibble
-   Explore the structure of the resulting list
-   Use the coefficients function to extract the regression coefficients
-   Use \$ to print out the R-Squared

```{r regression}

```
