---
title: 'Homework #17 Key'
author: "Lists and methods"
date: "Due: 30 March 2021"
output: 
  html_document:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
```

## Concepts

1.  Explain why a tibble is a list but not all lists are tibbles.

In R, vectors are one dimensional holders of information. Vectors have slots or positions in which information and be stored. The information can be referenced by specifying its slot number or name. Although not everyone agrees with Hadley Wickham, he states that there are two kinds of vectors:

-   atomic vectors
-   lists or recursive vectors

In most programming languages, lists are heterogeneous. This means that they can hold information of a variety of different classes: logical, integer, double, character, etc. The objects in a list can also have a variety of lengths. Lists can even contain other lists. Using lists within lists is a common practice.

A tibble is a collection of atomic vectors that are arranged in a list. Each of the atomic vectors must have the same length in a tibble. This means that tibbles are rectangular lists. A tibble is a special case of a list.

Because all lists do not contain just atomic vectors of the same length, lists in general do not satisfy the requirements for being a tibble.

2.  Explain how you use [], [[]], and \$ when subsetting a list as described in section 20.5.2.

We are able to extract information from lists using [], [[]], and \$.

The [] notation returns a sub-list and the result will always be a list.

The [[]] notation returns a simple component from a list.

The \$ approach extracts named elements of a list. It works similarly to [[]].

3.  Explain how class and method are related to each other. Use the summary method as an example to explain your answer.

Methods are a special kind of function. This generic function checks the class of an object that is being used as an argument to the function. Based on the class of the object, it chooses from among the family of functions included in the generic category. For example, the generic function or method summary can be applied to over 89 different classes of objects. If we ask R to summarize a data.frame, then it finds the specific function summary.data.frame. If we ask R to summarize an lm object, then it finds the specific function summary.lm to perform the operation. We can see the different methods for the generic summary call by using methods(summary)

```{r methods for summary}
methods(summary)
methods(summary) %>% length()
```

## Interpretation of code

Interpret the following R code chunk:

```{r extracting information from lists}
data_science <- list(transform = "Transform", model = "Model", visual = "Visual")
our_list <- list(one = TRUE, two = 3L, three = 500, four = "EDA", five = data_science)
our_list$five
our_list[1]
our_list[[1]]
our_list[["one"]]
our_list[1:2]
our_list$five$model
```

Line 1: Creates a name list with three slots. Slot one contains "Transform", two contains "Model", and three contains "Visual". The names of the slots are transform, model, and visual.

Line 2: Creates a second named list. The first one samed one is a logical atomic vector with the value TRUE. The second, named two, is an integer. Three and four are also atomic and are double and character. The fifth slot named five is the list created in line 1.

Line 3: Returns the data_science vector

Line 4: Returns a list which contains the atomic vector one

Line 5: Returns the atomic vector one

Line 6: Returns the same atomic vector as line 5

Line 7: Returns a list of one and two

Line 8: Returns the value "Model" as it goes to the slot named five and then gets the model slot from the resulting list.

## Write Code

Complete the following steps

-   Create a new object for a linear regression of cty on displ in the mpg tibble
-   Explore the structure of the resulting list
-   Use the coefficients function to extract the regression coefficients
-   Use \$ to print out the R-Squared

```{r regression}
regression_model <- lm(cty ~ displ, data = mpg)
regression_model %>% str()
regression_model %>% coefficients()
regression_model %>% summary() %>% str()
summary(regression_model)$r.squared
```
