---
title: 'Homework #17'
author: "Vectors"
date: "Due: 18 March 2020"
output: word_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
library(lubridate)
library(nycflights13)
```

## Concepts  

1.  Explain why a tibble is a list but not all lists are tibbles.  

2.  Explain how you use [], [[]], and $ when subsetting a list as described in section 20.5.2. 
    
## Interpretation of code  

Interpret the following R code chunk: 

Code Chunk #1

```{r code chunk for determining whether a temperature is above or below freezing}
data_science <- list(transform = "Transform", model = "Model", visual = "Visual")
our_list <- list(one = TRUE, two = 3L, three = 500, four = "EDA", five = data_science)
our_list$five
our_list[1]
our_list[[1]]
our_list[["one"]]
our_list[1:2]
our_list$five$model
```

Code Chunk #2

The nycflights13 library contains five tibbles: airlines, airports, flights, planes, and weather. The following code refers to these objects.

```{r code chunk interpretation, warning = FALSE, message = FALSE}
nyc <- list(airlines, airports, flights, planes, weather)
names(nyc) <- c("airlines", "airports", "flights", "planes", "weather")
nyc$planes %>% head(10)
nyc$flights$arr_delay %>% head(10)
```  

## Write Code

Complete the following steps  

*  Create a vector of the names of the members of your family
*  Create a vector of their birthdays as ymd (dates)
*  Create a tibble (with variable names) for your family from the vectors that you already created
*  Use dplyr to extract your name and birthday from your tibble

```{r family tibble}

```
