---
title: 'Homework #18'
author: "Time Series"
date: "Due: 1 April 2021"
output:
  html_document:
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(fpp3)
load(url("http://govfinance.byu.edu/govfinance/classes/dataViz/homework/HW10.RData"))
```

### Install the fpp3 library

### Concepts

Smoothing is a very important concept for data visualization. Explain how smoothing is affected in the following cases:

1.  binwidth and number of bins on histograms and freqpoly

2.  bandwidth (bw) on densities and violin plots

3.  span on loess smoothed lines

### Code Interpretation

#### Problem 1: Time series plot

```{r time series plot for ecommerce}
ecommerce %>%
  # as.Date is needed below to fix a knitting bug
  ggplot(aes(x = as.Date(date), y = ecommerce / 1000)) + 
  geom_point() + 
  geom_smooth(span = 0.3) +
  labs(title = "E-Commerce sales are growing exponentially",
       x = "",
       y = "Billions of Dollars")
```

#### Problem 2: STL decomposition

```{r stl decomposition}
ecommerce %>% 
  model(
    STL(ecommerce)
  ) %>% 
  components() %>% 
  autoplot()
```

Problem 3: ETS forecast

```{r ets forecast}
ecommerce %>% 
  model(
    ETS(ecommerce)
  ) %>% 
  forecast(h = 24) %>% 
  autoplot(ecommerce)
```

### Coding

#### Problem 1

Make a time series plot of Salt Lake City housing prices.

```{r problem 1}
# insert your code here
```

Problem 2

Use STL to look for the trend and seasonality SLC housing prices

```{r problem 2}
# insert your code here
```

Problem 3

Forecast SLC housing prices for the next two years

```{r problem 3}
# insert your code here
```
