---
title: "HW20"
author: "Ray Nelson"
date: "1 Apr 2019"
output:
  html_document:
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Concepts  

1.  Explain how you specify the equation that you want to estimate in a linear model.

2.  Explain how dependent and independent variables are related to response and explanatory (predictor) variables.

3.  Explain how you use categorical, numerical, and interactions in a linear model. What is their effect on the slopes and intercepts in your model.  
    
## Interpretation of code  

Interpret the following R code chunks: 

Code Chunk #1 Analysis of Variance Predictions

```{r ANOVA and prediction}
model_lm <- lm(mpg ~ am, data = mtcars)
model_lm %>% summary()
```

Code Chunk #2 Analysis of Covariance Predictions

```{r Analysis of Covariance I}
model_lm <- lm(mpg ~ wt + am + wt * am, data = mtcars)
model_lm %>% summary()
```

Code Chunk #3 Analysis of Covariance Predictions

```{r Analysis of Covariance II}
model_lm <- lm(mpg ~ wt + am, data = mtcars)
model_lm %>% summary()
```

## Write Code

Use the diamonds data set to create a model for predicting diamond prices. Use your model to predict the price of a specific diamond. Please come prepared to show your model to the class if you are selected as the guest lecturer.

```{r diamonds prediction model}

```