---
title: "HW22"
author: "Ray Nelson"
date: "8 Apr 2019"
output:
  html_document:
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Concepts  

Explain the idea of themes within ggplot.
    
## Interpretation of code  

Interpret the following R code chunks: 

Code Chunk #1 Zooming using coord_cartesian

```{r coord_cartesian zooming}
mpg %>% 
  ggplot(mapping = aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth() +
  coord_cartesian(xlim = c(5, 7), ylim = c(10, 30))
```

Code Chunk #2 Zooming using filtering

```{r filtered zooms}
mpg %>%
  filter(displ >= 5, displ <= 7, hwy >= 10, hwy <= 30) %>%
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth()
```  

Why are the loess lines in these two graphs different?

## Write Code

Problem 3 from 28.2.1 Exercises

Take an exploratory graphic that you've created during the semester, and add informative labels, annotations, scales, and themes to make it easier for others to understand.  

```{r diamonds prediction model}

```