---
title: "Gathering, Spreading, Separating, and Uniting"
author: "Ray Nelson"
date: "20 Feb 2019"
output:
  html_document:
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries
library(tidyverse)
library(fImport)
library(zoo)
library(Quandl)
```

## Different Tibble structures for the same data 

```{r alternative structures}
table1
table2
table3
table4a # cases
table4b # population
```

## Examples of the value of tidy data  

### Cases per 10,000 in population  

```{r cases relative to population}
table1 %>% 
  mutate(rate = cases / population * 10000)
```  

### Cases per year

```{r annual cases}
table1 %>% 
  count(year, wt = cases)
```  

### Visualization of changes over time  

```{r changes of time}
table1 %>% 
  ggplot(aes(x = year, y = cases)) +
    geom_line(aes(group = country), colour = "grey50", size = 1) +
    geom_point(aes(colour = country), size = 2) +
    labs(
      title = "Tuberculosis Cases",
      subtitle = "1999 and 2000",
      x = "",
      y = "Number of Cases"
    )
```  

## Gathering

### Turn Table 4a and 4b into one like Table 1  

```{r gather for table4}

# table 4a
table4a
table4a %>%
  gather(`1999`, `2000`, key = year, value = cases)

# table 4b
table4b
table4b %>% 
  gather(`1999`, `2000`, key = year, value = population)

# All the data together into a tidy tibble
tidy4a <- table4a %>%
  gather(`1999`, `2000`, key = year, value = cases)

tidy4b <- table4b %>%
  gather(`1999`, `2000`, key = year, value = population)

left_join(tidy4a, tidy4b)
```  

## Spread Table 2  

```{r spread table2, message = FALSE}
table2
cases <- table2[table2$type == "cases", ] %>% select(-type)
colnames(cases) <- c("country", "year", "cases")
population <- table2[table2$type == "population", ] %>% select(-type)
colnames(population) <- c("country", "year", "population")

left_join(cases, population)

# Same operation using spread

table2 %>%
  spread(key = type, value = count)
```


## Example of how time series gathering  

### Use FRED as the data source 

#### Load functions for time series calculations

```{r ts manipulation functions}
# Get time series functions
"http://govfinance.byu.edu/govfinance/quantTools/R/manipulations.R" %>%
  url() %>%
  source()
```  

#### Get Data from FRED

```{r FRED}
# Time series of income and sales taxes in Utah using from FRED using fImport
revenue <- fredSeries(c("UTSLGRTAX", "UTINCTAX"), from = "1992-01-01") %>%
  as.zoo() %>% 
  pca()
colnames(revenue) <- c("sales", "income")
revenue
revenue %>% str()
```

#### Create tibbles and data frames from the FRED data  

```{r create tibbles and data frames from FRED data}
# Extract the date from the zoo object
dates <- revenue %>%
  time() %>%
  as.Date()

# Create a tibble
revenue_tbl <- tibble(dates, as.vector(revenue$sales), as.vector(revenue$income))
colnames(revenue_tbl) <- c("dates", "sales", "income")
revenue_tbl

# Create a data frame
revenue_df <- data.frame(dates, revenue)
colnames(revenue_df) <- c("dates", "sales", "income")
revenue_df
```  

### Use Quandl to accomplish the same thing

```{r create tibbles and data frames with Quandl data}
# Time series of income and sales taxes in Utah using Quandl
revenue <- Quandl(c("FRED/UTSLGRTAX", "FRED/UTINCTAX"),
                  transform = "rdiff") %>% as.tibble()
colnames(revenue) <- c("dates", "sales", "income")
revenue_tbl <- revenue %>% mutate(sales = sales * 100, income = income * 100)
revenue_tbl %>% str()
revenue_tbl
```  

### Time series graphs sales an income taxes by themselves  

```{r sales and income by themselves}
# Graph of sales by itself
revenue_tbl %>% 
  ggplot(aes(x = dates, y = sales)) +
    geom_line() +
    geom_hline(yintercept = 0, color = "red") +
    labs(title = "Utah Sales Tax Growth Rate",
         subtitle = "1992 - 2016",
         x = "",
         y = "Percentage Change")

revenue_tbl %>% 
  ggplot(aes(x = dates, y = income)) +
  geom_line() +
  geom_hline(yintercept = 0, color = "red")  +
  labs(title = "Utah Income Tax Growth Rate",
       subtitle = "1992 - 2016",
       x = "",
       y = "Percentage Change")
```  

### Use graphs to compare sales and income  

```{r gather sales and revenue}  
# Gather the data together
plot_data <- revenue_tbl %>% 
  gather(sales, income, key = "tax", value = "revenue")
```  

```{r comparison of sales and income}
# Time series plot of both in the same graph
plot_data %>% 
  ggplot(aes(x = dates, y = revenue, color = tax)) +
    geom_line() +
    geom_hline(yintercept = 0, color = "red") +
    labs(title = "Utah Sales and Income Tax Growth Rates",
         subtitle = "1992 - 2016",
         x = "",
         y = "Percentage Change")

# Time series plot of each in its own panel
plot_data %>% 
  ggplot(aes(x = dates, y = revenue)) +
    geom_line(aes(color = tax), show.legend = FALSE) +
    facet_grid(rows = vars(tax)) +
    geom_hline(yintercept = 0, color = "red", alpha = 0.25) +
    labs(title = "Utah Sales and Income Tax Growth Rates",
         subtitle = "1992 - 2016",
         x = "",
         y = "Percentage Change")
```  

### Violin plots to compare the distribution of sales and income tax revenues  

```{r violin plot comparison}
# Violin Plot
plot_data %>%
   ggplot(aes(x = tax, y = revenue)) +
    geom_violin(fill = "lightgreen") +
    geom_boxplot(width = 0.25, fill = "lightblue") +
    labs(title = "Comparison of Sales and Income Tax Growth Rates",
         subtitle = "State of Utah, 1992 - 2016",
      x = "Tax",
      y = "Percentage Change") +
    coord_flip()
```  








