---
title: "Homework #20"
author: "Shiny I"
date: "Due: 6 April 2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
```

## Concepts  

Explain the function and interaction of the ui and server parts of a shiny app.
    
## Interpretation of code  

Code Chunk #1 UI part of a shiny app

```{r ui}
ui <- fluidPage(
    titlePanel("Old Faithful Geyser Data"),

    sidebarLayout(
        sidebarPanel(
            sliderInput("bins",
                        "Number of bins:",
                        min = 1,
                        max = 50,
                        value = 30)
        ),

        mainPanel(
           plotOutput("distPlot")
        )
    )
)
```

Code Chunk #2 Server Part of a Shiny app

```{r server}
server <- function(input, output) {

    output$distPlot <- renderPlot({

        x    <- faithful[, 2]
        bins <- seq(min(x), max(x), length.out = input$bins + 1)

        hist(x, breaks = bins, col = 'darkgray', border = 'white')
    })
}
```  


## Write Code

Alter the basic shiny app so that it draws a histogram with number of bins that are determined by the user for the hwy variable in the mpg tibble. Attach your R script file to your homework.
