---
title: "Homework #20"
author: "Formulas and Model Families"
date: "Due: 1 April 2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
```

## Concepts  

1.  Explain how you specify the equation that you want to estimate in a linear model.

2.  Explain how dependent and independent variables are related to response and explanatory (predictor) variables.
    
## Interpretation of code  

Code Chunk #1 Analysis of Variance Predictions

```{r ANOVA and prediction}
model_lm <- lm(mpg ~ am, data = mtcars)
model_lm %>% summary()
```

Code Chunk #2 Analysis of Covariance Predictions

```{r Analysis of Covariance I}
model_lm <- lm(mpg ~ wt + am + wt * am, data = mtcars)
model_lm %>% summary()
```

Code Chunk #3 Analysis of Covariance Predictions

```{r Analysis of Covariance II}
model_lm <- lm(mpg ~ wt + am, data = mtcars)
model_lm %>% summary()
```

## Write Code

Use the diamonds data set to create a model for predicting diamond prices. Use your model to predict the price of a specific diamond. Please come prepared to show your model to the class if you are selected as the guest lecturer.

```{r diamonds prediction model}

```