---
title: "Homework #22"
author: "Geospatial"
date: "Due: 13 April 2020"
output:
   html_document:
     code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
library(leaflet)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
```

## Concepts  

1. Define projections and layers within the context of geospatial data visualization. Why are they so important? 
2. Compare and contrast choropleths with cartograms.
    
## Interpretation of code  

### Code chunk #1

```{r First code chunk}
latitude <- 40.2504
longitude <- -111.6525

(TNRB <- leaflet() %>% 
    addProviderTiles(provider = providers$CartoDB.Positron) %>% 
    addMarkers(lng = longitude,
               lat = latitude,
               popup = "Marriott School of Business")
)
```

### Code chunk #2

```{r Second code chunk}
world <- ne_countries(scale = "medium", returnclass = "sf")

(world %>% 
  ggplot() +
  geom_sf() +
  coord_sf(xlim = c(-88, -78),
           ylim = c(24.5, 33),
           expand = FALSE)
)
```

## Write Code

Write a function that locates and labels a point on a map.

* The function should take longitude, latitude, and a label as its arguments.
* The function should use maps provided by Stamen.Toner.
* Use your function to draw a map with the location of your home or apartment.
