---
title: "Factors"
author: "MPA 634: Data Science for Managers"
date: "26 Feb 2020"
output:
  html_document:
    code_folding: "hide"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
# Load libraries
library(tidyverse)
library(forcats)
library(kableExtra)
library(ggmosaic)
```

## Value of factors in modeling

```{r linear model with factors}
mpg %>% summary()
lm(hwy ~ drv + displ + drv * displ, data = mpg) %>% 
  summary()

mpg$drv <- as.factor(mpg$drv)
mpg %>% summary()
lm(hwy ~ drv + displ + drv * displ, data = mpg) %>% 
  summary()
```

## 15.2 Creating factors

### Create two initial vectors, one of which has an error.

```{r initial vectors}
(x1 <- c("Dec", "Apr", "Jan", "Mar"))
(x2 <- c("Dec", "Apr", "Jam", "Mar"))

sort(x1)
class(x1)
```

### Create a vector of labels and then create two new factors

```{r create a vector of label}
month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

(y1 <- factor(x1, levels = month_levels))
(y2 <- factor(x2, levels = month_levels))
sort(y1)
class(y1)
```

### Order of levels  

#### Alphabetical order

```{r levels in alphabetical order}
factor(x1)
```  

#### Order of the original data

```{r levels in data order}
unique(x1)
(f1 <- factor(x1, levels = unique(x1)))

(f2 <-factor(x1) %>% fct_inorder())

f2 %>% levels()
```  

## 15.3 General Social Survey

### Familiarity with the data  

```{r familiarity with data}
? gss_cat
gss_cat %>% summary
```  

### Count the frequency in a tibble

```{r factors in a tibble}
gss_cat %>% 
  count(race)
```  

### Visual representation of a factor using geom_bar

```{r frequency with geom_bar}
gss_cat$race %>% levels()
gss_cat %>% 
  ggplot(aes(x = race, fill = race)) +
  geom_bar(show.legend = FALSE) +
  labs(title = "Implicit count from within geom_bar")
```

### Counting multiple factors or traditional cross tabulations

#### Simple count using count from dplyr

```{r simple counts}
gss_cat %>% 
  count(race, marital)
```

#### Spread the results of count

```{r spread and kable}
gss_cat %>% 
  count(race, marital) %>% 
  spread(marital, n) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

#### Mosaic plot of cross tab

```{r mosaic chart}
gss_cat %>% 
  ggplot() +
  geom_mosaic(aes(x = product(race), fill = fct_infreq(marital)),
          na.rm = TRUE,
          show.legend = FALSE)
```

## 15.4 Modifying factor order

### TV hours by religion

#### Original Order

```{r TV hours by religion}
# Create the relig_summary tibble using summary to calculate the mean and tv hours
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

relig_summary %>%
  ggplot(aes(tvhours, relig)) +
  geom_point() +
  labs(title = "Initial order of factors",
       x = "Hours",
       y = "")
```

#### Religion ordered by tvhours

```{r religion sorted by tvhours}
relig_summary %>%
  ggplot(aes(tvhours, fct_reorder(relig, tvhours))) +
  geom_point() +
  labs(title = "Initial order of factors",
       x = "Hours",
       y = "")
```

### Age by marital status in original order

```{r age by marital status}
# Calculate the proportion in each age and marital status category
by_age <- gss_cat %>%
  filter(!is.na(age)) %>%
  count(age, marital) %>%
  group_by(age) %>%
  mutate(prop = n / sum(n))

by_age %>%
  ggplot(aes(x = age, y = prop, color = marital)) +
  geom_line(na.rm = TRUE) +
  labs(title = "Original ordering of factors")
```

### Age by marital status with legend that matches the last observations

```{r altered legend}
by_age %>%
  ggplot(aes(age, prop, color = fct_reorder2(marital, age, prop))) +
  geom_line(na.rm = TRUE) +
  labs(title = "Lines order to match the legend order",
       x = "Age",
       y = "Percentage",
    colour = "marital")
```

### Pareto chart of marital status

```{r pareto chart of marital status}
gss_cat %>%
  mutate(marital = marital %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(x = marital, fill = marital)) +
  geom_bar(show.legend = FALSE) +
  labs(
    title = "Bar chart with factors ordered by frequency",
    subtitle = "This creates a Pareto Chart.",
    x = "",
    y = "Number of Respondents"
  )
```

## 15.5 Modifying factor levels

### Initial levels for party id

```{r modifying factor levels}
# Original factor levels and labels
gss_cat %>% count(partyid)
``` 

### Change the labels for some levels of a factor

```{r change labels for some levels}
gss_cat %>%
  mutate(
    partyid = fct_recode(
      partyid,
      "Republican, strong"    = "Strong republican",
      "Republican, weak"      = "Not str republican",
      "Independent, near rep" = "Ind, near rep",
      "Independent, near dem" = "Ind, near dem",
      "Democrat, weak"        = "Not str democrat",
      "Democrat, strong"      = "Strong democrat"
    )
  ) %>%
  count(partyid)
```

### Combine multiple factors levels into a single category  

```{r combine multiple levels into a single category}
gss_cat %>%
  mutate(
    partyid = fct_recode(
      partyid,
      "Republican, strong"    = "Strong republican",
      "Republican, weak"      = "Not str republican",
      "Independent, near rep" = "Ind,near rep",
      "Independent, near dem" = "Ind,near dem",
      "Democrat, weak"        = "Not str democrat",
      "Democrat, strong"      = "Strong democrat",
      "Other"                 = "No answer",
      "Other"                 = "Don't know",
      "Other"                 = "Other party"
    )
  ) %>%
  count(partyid)
```

### Another example of assigning multiple categories to the same category
### Uses fct_collapse

```{r factor collapse}
gss_cat %>%
  mutate(partyid = fct_collapse(
    partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  )) %>%
  ggplot(aes(x = partyid)) +
  geom_bar(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Independents are the largest group of respondents",
    x = "",
    y = "Number"
  )
```  
