---
title: "Factors"
author: "Ray Nelson"
date: "26 Feb 2019"
output:
  html_document:
    code_folding: "hide"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
# Load libraries
library(tidyverse)
library(forcats)
library(janitor)
# 
library(kableExtra)
```

## 15.2 Creating factors

### Create two initial vectors, one of which has an error.

```{r initial vectors}
(x1 <- c("Dec", "Apr", "Jan", "Mar"))
(x2 <- c("Dec", "Apr", "Jam", "Mar"))

sort(x1)
class(x1)
```

### Create a vector of labels and then create two new factors

```{r create a vector of label}
month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

(y1 <- factor(x1, levels = month_levels))
(y2 <- factor(x2, levels = month_levels))
sort(y1)
class(y1)
```

### Order of levels  

#### Alphabetical order

```{r levels in alphabetical order}
# levels in alphabetical order
factor(x1)
```  

#### Order of the data

```{r levels in data order}
# Order of the original data
unique(x1)
(f1 <- factor(x1, levels = unique(x1)))

(f2 <-factor(x1) %>% fct_inorder())

f2 %>% levels()
```  

## 15.3 General Social Survey

### Familiarity with the data  

```{r familiarity with data}
? gss_cat
data(gss_cat)
gss_cat %>%
  View()
```  

### Value of factors in a tibble

```{r factors in a tibble}
# dplyr count
gss_cat %>% 
  count(race)

# ggplot geom_bar with implied count
gss_cat %>%
  ggplot(aes(x = race, fill = race)) +
  geom_bar(show.legend = FALSE) +
  labs(title = "Implicit count from within geom_bar")

# ggplot geom_bar that shows the missing values
gss_cat %>% 
  ggplot(aes(x = race, fill = race)) +
  geom_bar(show.legend = FALSE) +
  scale_x_discrete(drop = FALSE) +
  labs(
    title = "Include the category of missing values in the bar chart"
  )
```  

### Counting multiple factors or traditional cross tabulations

#### Simple count using count from dplyr

```{r simple counts}
gss_cat %>% 
  count(race, marital)
```

```{r spread and kable}
gss_cat %>% 
  group_by(race, marital) %>% 
  summarise(n = n()) %>% 
  spread(marital, n) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

```{r tabyl and kable}
gss_cat %>% 
  tabyl(race, marital) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## 15.4 Modifying factor order

### TV hours by religion

```{r TV hours by religion}

# Create the relig_summary tibble using summary to calculate the mean and tv hours
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

relig_summary$relig %>%
  levels()

relig_summary %>%
  ggplot(aes(tvhours, relig)) +
  geom_point() +
  labs(title = "Initial order of factors",
       x = "Hours",
       y = "")
```

### Reordered factors for a more meaningful graph

```{r reordered factors}
# Largest to smallest
relig_summary %>%
  ggplot(aes(tvhours, fct_reorder(relig, tvhours, .desc = TRUE))) +
  geom_point() +
  labs(title = "Largest to smallest",
       x = "Hours",
       y = "")

# Smallest to largest
relig_summary %>%
  mutate(relig = fct_reorder(relig, tvhours)) %>%
  ggplot(aes(tvhours, relig)) +
  geom_point() +
  labs(title = "Smallest to largest",
       x = "Hours",
       y = "") 
```  

### Reorder the factors in the tibble
```{r reorder factors within the tibble}
relig_summary$relig %>% 
  levels()
  
relig_summary <- relig_summary %>%
  mutate(relig = fct_reorder(relig, tvhours, desc = TRUE))

relig_summary$relig %>% 
  levels()
```

### Average age for each reported income levels

```{r average age for different income levels}
# Initial summary calculations 
(rincome_summary <- gss_cat %>%
   group_by(rincome) %>%
   summarise(age = mean(age, na.rm = TRUE),
             n = n()))

# Graph with income ordered by age
rincome_summary %>%
  mutate(rincome = fct_reorder(rincome, age))  %>%
  ggplot(aes(x = age, y = rincome)) +
  geom_point() +
  labs(title = "Income ordered by age")

# Graph with income 
rincome_summary %>%
  ggplot(aes(x = age, y = rincome)) +
  geom_point() +
  labs(title = "No ordering")
```

### Turning income into an ordinal factor

```{r current order of income}
levels(gss_cat$rincome)
```

```{r convert to ordinal}
income_levels <- c("Lt $1000", "$1000 to 2999", "$3000 to 3999", "$4000 to 4999",
                   "$5000 to 5999", "$6000 to 6999", "$7000 to 7999", "$8000 to 9999",
                   "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 or more",
                   "No answer", "Don't know", "Refused", "Not applicable")
gss_cat$rincome <- factor(gss_cat$rincome, levels = income_levels, ordered = TRUE)
gss_cat$rincome %>% levels()
```

```{r graph ordered by income}
gss_cat %>%
   group_by(rincome) %>%
   summarise(age = mean(age, na.rm = TRUE),
             n = n()) %>% 
  ggplot(aes(x = age, y = rincome)) +
  geom_point() +
  labs(title = "Income as an ordered factor")
```

### Age by marital status

```{r age by marital status}
# Calculate the proportion in each age and marital status category
by_age <- gss_cat %>%
  filter(!is.na(age)) %>%
  count(age, marital) %>%
  group_by(age) %>%
  mutate(prop = n / sum(n))

by_age %>%
  ggplot(aes(x = age, y = prop, color = marital)) +
  geom_line(na.rm = TRUE) +
  labs(title = "Original ordering of factors")

by_age %>%
  ggplot(aes(age, prop, color = fct_reorder2(marital, age, prop))) +
  geom_line(na.rm = TRUE) +
  labs(title = "Lines order to match the legend order",
       x = "Age",
       y = "Percentage",
    colour = "marital")

gss_cat %>%
  mutate(marital = marital %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(x = marital, fill = marital)) +
  geom_bar(show.legend = FALSE) +
  labs(
    title = "Bar chart with factors ordered by frequency",
    subtitle = "This creates a Pareto Chart.",
    x = "",
    y = "Number of Respondents"
  )
```


## 15.5 Modifying factor levels

### Initial levels for party ide

```{r modifying factor levels}
# Original factor levels and labels
gss_cat %>% count(partyid)

``` 

### Change the labels for some levels of a factor

```{r change labels for some levels}
gss_cat %>%
  mutate(
    partyid = fct_recode(
      partyid,
      "Republican, strong"    = "Strong republican",
      "Republican, weak"      = "Not str republican",
      "Independent, near rep" = "Ind, near rep",
      "Independent, near dem" = "Ind, near dem",
      "Democrat, weak"        = "Not str democrat",
      "Democrat, strong"      = "Strong democrat"
    )
  ) %>%
  count(partyid)
```

### Combine multiple factors levels into a single category  

```{r combine multiple levels into a single category}
gss_cat %>%
  mutate(
    partyid = fct_recode(
      partyid,
      "Republican, strong"    = "Strong republican",
      "Republican, weak"      = "Not str republican",
      "Independent, near rep" = "Ind,near rep",
      "Independent, near dem" = "Ind,near dem",
      "Democrat, weak"        = "Not str democrat",
      "Democrat, strong"      = "Strong democrat",
      "Other"                 = "No answer",
      "Other"                 = "Don't know",
      "Other"                 = "Other party"
    )
  ) %>%
  count(partyid)
```

### Another example of assigning multiple categories to the same category
### Uses fct_collapse

```{r factor collapse}
gss_cat %>%
  mutate(partyid = fct_collapse(
    partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  )) %>%
  ggplot(aes(x = partyid)) +
  geom_bar(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Independents are the largest group of respondents",
    x = "",
    y = "Number"
  )
```  
