---
title: "Final Review PowerPoint"
author: "Ray Nelson"
date: "12/12/2018"
output: powerpoint_presentation
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readxl)
library(gridExtra)
```

## Description of Data

```{r read in the data}
companies <- read_excel("Companies.xlsx")
```  

The data set Companies.xlsx contains the following variables:

- type: Type of company, Computer or Pharmaceutical
- size: Categorical variable small or big
- sales: In millions of dollars
- profits: In millions of dollars
- employees: Numeric
- profit_per_employee: Numeric
- assets: In millions of dollars
- profit_sales_percent: Numeric

## Variables in the data set

```{r variables in companies tibble}
companies %>% glimpse()
```  

### Relationship between sales and profits linear?

The linear model suggests that there is a strong relationship between sales and profits but there is one very large company that is exerting undue influence on the estimated equation.

```{r sales and profits scatterplot, message = FALSE}
companies %>% 
  ggplot(aes(x = sales, y = profits)) +
    geom_point() +
    geom_smooth(color = "hotpink") +
    geom_smooth(method = "lm", color = "lightblue")
```  

## The linear model

The linear model has a large R-Squared.  

```{r linear model for all data}
lm(profits ~ sales, data = companies) %>% 
  summary()
``` 

## Let's exclude the very large company from our analysis  

```{r remove large company from scatterplot, message = FALSE}
companies %>% 
  filter(sales < max(sales)) %>% 
  ggplot(aes(x = sales, y = profits)) +
    geom_point() +
    geom_smooth(color = "hotpink") +
    geom_smooth(method = "lm", color = "lightblue")
```

## Model without the large company

```{r linear model without the large company}
lm(profits ~ sales, data = companies %>% filter(sales < max(sales))) %>%
  summary()
```  

## Distribution of profit with the large company   

There is one company with really large profits in absolute terms.

```{r large company effect on distribution}
companies %>%
  ggplot(aes(x = profits)) +
    geom_density(fill = "lightblue")
```

## Distribution of profit per employee  

The large company also has large sales. What happens if we look at profit relative to employees.  

```{r sales relative to employees}
companies %>%
  ggplot(aes(x = profit_per_employee)) +
    geom_density(fill = "lightblue")
```

## Distribution of profit as a percentage of sales  

The large company also has large sales. What happens if we look at profit relative to sales.

```{r sales as a percentage of profits}
companies %>%
  ggplot(aes(x = profit_sales_percent)) +
    geom_density(fill = "lightblue")
```

## Control Distribution of Sales with Log Transformation   

The large company also has large sales. What happens to the distribution if we take the natural log of sales?  

```{r comparison of sales}
untransformed <- companies %>%
  ggplot(aes(x = sales)) +
    geom_density(fill = "lightblue") +
    labs(
      title = "(a) Untransformed Data"
    )
transformed <- companies %>%
  ggplot(aes(x = log(sales))) +
    geom_density(fill = "lightblue") +
    labs(
      title = "(b) Transformed Data"
    )
grid.arrange(untransformed, transformed, nrow = 1)
```


