---
title: "Midterm I Winter 2020"
author: "Data Science for Managers"
date: "Due by class: 2/18/2020"
output:
  html_document:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

Submit an R Markdown for the following two problems. In order to get full credit, your R Markdown document must knit properly.  

## Problem 1

Please write code that completes the following:

*  Create a new variable for the departure hour from the scheduled departure time variable
*  Calculate the number of flights that leave during each hour and each airport
*  Draw a bar chart for each of the three New York airports
*  Color the bars based on their origin airport
*  It is not required that your graph replicate the graph labels. We will cover that in Chapter 28.


```{r problem 1}
flights %>% 
  mutate(
    dep_hour = sched_dep_time %/% 100
  ) %>% 
  group_by(origin, dep_hour) %>%
  summarise(
    number_flights = n()
  ) %>% 
  ggplot(aes(dep_hour, number_flights, fill = origin)) +
  geom_col(show.legend = FALSE) +
  facet_grid(rows = vars(origin)) +
  labs(
    title = "Number of flights leaving each New York airport by departure hour",
    subtitle = "Departure hour calculated from the scheduled departure time",
    x = "Departure hour",
    y = "Total number of flights\n"
  )
```

## Problem 2

Please write code that completes the following:

*  Determine the number of destinations for each NYC airport
*  Create a bar chart that depicts the number of destinations
*  Color the bars based on the departure airport
*  Order the bars from smallest to largest number of destinations (This is called a Pareto Chart)
*  It is not required that your graph replicate the graph labels. We will cover that in Chapter 28.

```{r Problem 2}
flights %>% 
  group_by(origin) %>% 
  summarise(
    destinations = n_distinct(dest)
  ) %>% 
  ggplot(aes(x = reorder(origin, destinations, FUN = sum), y = destinations)) +
  geom_col(aes(fill = origin), show.legend = FALSE) +
  labs(
    title = "Number of distinations for each New York airport",
    subtitle = "Sorted by number of destinations",
    x = "New York Airport",
    y = "Number of Destinations"
  )
```

