---
title: "R Markdown Exercise"
author: "MPA 634: Data Science for Managers"
date: "12 Jan 2020"
output:
  html_document:
    code_folding: hide
editor_options: 
  markdown: 
    wrap: 80
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Data Science for Managers

The intersection of statistics, computer science, and domain knowledge defines
data science. The process whereby data scientists explore, wrangle, program,
model, and communicate quantitative relationships is greatly facilitated by the
statistical computer language R.

## Learning Outcomes

1.  Explore data through visualization and transformation
2.  Organize data into constructs that are useful for visualizing and modeling
3.  Master the philosophy, basics, and components of the R language
4.  Communicate clearly the insights gleaned from data exploration

## Distribution of a Random Variable

### Definition

1.  Range of values
2.  Frequencies or probabilities

## Characteristics

-   **Location or center** (*mean and median*)
-   **Scale or spread** (*standard deviation and interquartile range*)
-   **Symmetry or skewness** (*skewness coefficient*)
-   **Outliers, kurtosis, or tail thickness** (*coefficient of excess*)

## Categorical Variables (Nominal and Ordinal)

Bar charts help us visualized the distribution of a categorical variable.

```{r bar chart}
ggplot(data = mpg,
       mapping = aes(x = drv)) +
  geom_bar() +
  coord_flip()
```

## Numerical Variables

### Histogram

```{r histogram}
ggplot(data = mpg,
       mapping = aes(x = hwy)) +
  geom_histogram(bins = 15)
```

Determining the bin width or number of bins is an important decision when
drawing a histogram.

### Density Plot

```{r density}
ggplot(data = mpg,
       aes(x = hwy)) +
  geom_density()
```

### Boxplot

```{r boxplot}
ggplot(data = mpg,
       aes(x = factor(0), y = hwy)) +
  geom_boxplot() +
  coord_flip()
```

The boxplot helps us judge the location, scale, symmetry, and outliers of a
distribution.

-   Location: center line in the boxplot is the median.

-   Scale: length of the rectangle is the interquartile range.

-   Symmetry:

    -   Position of the median in the box
    -   Relative length of the whiskers
    -   Number of outliers in each tail

-   Kurtosis: number of outliers

### Violin Plot

Reflecting a density trace along a horizontal axis gives a violin plot.

```{r violin}
ggplot(data = mpg,
       aes(x = hwy, y = factor(0))) +
  geom_violin()
```

### True Violin Plot

A true violin plot is a composite of a boxplot and a reflected density trace. 

```{r true violin}
ggplot(data = mpg,
       mapping = aes(x = hwy, y = factor(0))) +
  geom_violin(fill = "lightblue") +
  geom_boxplot(width = 0.10, fill = "grey60") +
  stat_summary(
    fun = "mean",
    geom = "point",
    shape = 23,
    size = 1.5,
    fill = "white"
  )
```
