---
title: "R Markdown Exercise"
author: 'MPA 634: Data Science for Managers'
date: "9/13/2019"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Distribution of a Random Variable

## Definition

1.  Range of values
1.  Frequencies or probabilities

## Characteristics

*  **Location or center** (*mean and median*)
*  **Scale or spread** (*standard deviation and interquartile range*)
*  **Symmetry or skewness** (*skewness coefficient*)
*  **Outliers, kurtosis, or tail thickness** (*coefficient of excess*)

## Categorical Variables (Nominal and Ordinal)

Bar charts help us visualized the distribution of a categorical variable.

```{r bar chart}
ggplot(data = mpg, mapping = aes(x = drv)) +
  geom_bar() +
  coord_flip()
```

## Numerical Variables

### Histogram

```{r histogram}
ggplot(data = mpg, mapping = aes(x = hwy)) +
  geom_histogram(bins = 15)
```

Determining the bin width or number of bins is an important decision when drawing a histogram.

### Density Plot

```{r density}
ggplot(data = mpg, aes(x = hwy)) +
  geom_density()
```

### Boxplot

```{r boxplot}
ggplot(data = mpg, aes(x = factor(0), y = hwy)) +
  geom_boxplot() +
  coord_flip()
```

The boxplot helps us judge the location, scale, symmetry, and outliers of a distribution.

*  Location: center line in the boxplot is the median.
*  Scale: length of the rectangle is the interquartile range.
*  Symmetry:
   +  Position of the median in the box
   +  Relative length of the whiskers
   +  Number of outliers in each tail
*  Kurtosis: number of outliers

### Violin Plot

```{r violin}
ggplot(data = mpg, aes(x = factor(0), y = hwy)) +
  geom_violin() +
  coord_flip()
```

### True Violin Plot

```{r true violin}
ggplot(data = mpg, aes(x = factor(0), y = hwy)) +
  geom_violin() +
  geom_boxplot(width = 0.25) +
  coord_flip()
```
