---
title: "Relational Data"
author: "MPA 634: Data Science for Managers"
date: "24 Feb 2020"
output:
  html_document:
    code_folding: "hide"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
# Load libraries
library(tidyverse)
library(nycflights13)
```

## Overview of tables to be manipulated  

### Flights Table

```{r flights tibble}
flights %>% glimpse()
```

### Airlines Table

```{r airlines tibble}
airlines %>% glimpse()
```  

### Airports Table

```{r airports tibble}
airports %>% glimpse()
```  
### Planes Table  

```{r planes tibble}
planes %>% glimpse()
```  
### Weather Table  

```{r weather tibble}
weather %>% glimpse()  
```  
## Keys: Primary and Foreign  

### Check to see if keys are unique for planes tibbles

```{r check planes}
planes %>% 
  count(tailnum) %>% 
  filter(n > 1)
```  
### Check to see if keys are unique for planes tibbles

```{r check weather}
weather %>% 
  count(year, month, day, hour, origin) %>% 
  filter(n > 1)
```

### Sometimes we don't have an explicit primary key 

#### Check uniqueness of flight numbers

```{r unique flights}  
flights %>% 
  count(year, month, day, flight) %>% 
  filter(n > 1)
```  

#### Check uniquess of tail number

```{r unique tail number}  
flights %>% 
  count(year, month, day, tailnum) %>% 
  filter(n > 1)
``` 

## Mutating joins  

### Subset of columns from flights

```{r abbreviated flights}
(flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
)
```  

### Add the full airline name to the flights2 data.  

```{r full airline name}
flights2 %>%
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")
```  

### Understanding joins  

#### Create the two tibbles x and y  

```{r join tibbles}
(x <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2",
  3, "x3")
)

(y <- tribble(
  ~key, ~val_y,
  1, "y1",
  2, "y2",
  4, "y3")
)
```  

### inner join
```{r inner join}
x %>% 
  inner_join(y, by = "key")
```  

### outer join

#### left join
```{r left join}
x %>% 
  left_join(y, by = "key")
```  

#### right join
```{r right join}
x %>% 
  right_join(y, by = "key")
```  

#### full join
```{r full join}
x %>% 
  full_join(y, by = "key")
```  

### Duplicate Keys

#### Two tibbles with duplicate keys

```{r duplicate}
(x <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2",
  2, "x3",
  1, "x4")
)

(y <- tribble(
  ~key, ~val_y,
  1, "y1",
  2, "y2")
)
``` 

#### Left join

```{r left join duplicate keys}
x %>%
  left_join(y, by = "key")
```

### Defining the key columns  

```{r default key}
flights2 %>% 
  left_join(weather)
```  

### Using the by  

```{r use by}
flights2 %>% 
  left_join(planes, by = "tailnum")
```  

### Using the by with two different tables with different key names 

#### Variables in flights2

```{r flights2 variables}
flights2 %>%  glimpse()
```
#### Variables in airports
```{r airports variables}
airports %>%  glimpse()
```

#### Left join for destination

```{r different key names destination}
flights2 %>% 
  left_join(airports, c("dest" = "faa"))
```  

#### Left join for origin

```{r different key names origin}
flights2 %>% 
  left_join(airports, c("origin" = "faa"))
```  

### Filtering joins 

#### Find the top 10 destinations

```{r top ten destinations}
top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest  
```  

#### Choose all flights that go to the top ten destinations using filter

```{r top ten filter}
flights %>% 
  filter(dest %in% top_dest$dest)
```  

#### Choose all flights that go to the top ten destinations using semi_join
```{r top ten semi join}
flights %>% 
  semi_join(top_dest)
```  

#### Choose all flights that do not go to the top ten destinations using anti_join 

```{r not top ten}
flights %>%
  anti_join(top_dest)
``` 