---
title: "Relational Data"
author: "MPA 634: Data Science for Managers"
date: "2 March 2021"
output:
  html_document:
    code_folding: "hide"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
# Load libraries
library(tidyverse)
library(nycflights13)
data(flights)
data(planes)
data(weather)
data(airlines)
data(airports)
```

## Overview of tables to be manipulated

### Flights Table

```{r flights tibble}
flights %>% glimpse()
```

### Airlines Table

```{r airlines tibble}
airlines %>% glimpse()
```

### Airports Table

```{r airports tibble}
airports %>% glimpse()
```

### Planes Table

```{r planes tibble}
planes %>% glimpse()
```

### Weather Table

```{r weather tibble}
weather %>% glimpse()  
```

## Keys: Primary and Foreign

### Check to see if keys are unique for planes tibbles

```{r check planes}
planes %>% 
  count(tailnum) %>% 
  filter(n > 1)
```

#### Check to see if keys are unique for each month, day, hour, and origin airport in the weather tibble

```{r check weather}
weather %>% 
  count(year, month, day, hour, origin) %>% 
  filter(n > 1)
```

### Sometimes we don't have an explicit primary key

#### Check uniqueness of flight numbers

```{r unique flights}
flights %>% 
  count(year, month, day, flight) %>% 
  filter(n > 1)
```

#### Check uniqueness of tail number on each day

```{r unique tail number}
flights %>% 
  count(year, month, day, tailnum) %>% 
  filter(n > 1)
```

## Mutating joins

### Subset of columns from flights

```{r abbreviated flights}
(flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
)
```

### Add the full airline name to the flights2 data.

```{r full airline name}
flights2 %>%
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")
```

### Understanding joins

#### Create the two tibbles x and y

```{r join tibbles}
(x <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2",
  3, "x3")
)

(y <- tribble(
  ~key, ~val_y,
  1, "y1",
  2, "y2",
  4, "y3")
)
```

### inner join

```{r inner join}
x %>% 
  inner_join(y, by = "key")
```

### outer join

#### left join

```{r left join}
x %>% 
  left_join(y, by = "key")
```

#### right join

```{r right join}
x %>% 
  right_join(y, by = "key")
```

#### full join

```{r full join}
x %>% 
  full_join(y, by = "key")
```

### Duplicate Keys

#### Two tibbles with duplicate keys

```{r duplicate}
(x <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2",
  2, "x3",
  1, "x4")
)

(y <- tribble(
  ~key, ~val_y,
  1, "y1",
  2, "y2")
)
```

#### Left join

```{r left join duplicate keys}
x %>%
  left_join(y, by = "key")
```

### Defining the key columns

```{r default key}
flights2 %>% 
  left_join(weather)
```

### Using the by

```{r use by}
flights2 %>% 
  left_join(planes, by = "tailnum")
```

### Using the by with two different tables with different key names

#### Variables in flights2

```{r flights2 variables}
flights2 %>%  glimpse()
```

#### Variables in airports

```{r airports variables}
airports %>%  glimpse()
```

#### Left join for destination

```{r different key names destination}
flights2 %>% 
  left_join(airports, c("dest" = "faa"))
```

#### Left join for origin

```{r different key names origin}
flights2 %>% 
  left_join(airports, c("origin" = "faa"))
```

### Filtering joins

#### Find the top 10 destinations

```{r top ten destinations}
(top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  slice_head(n = 10)
)
```

#### Choose all flights that go to the top ten destinations using filter

```{r top ten filter}
flights %>% 
  filter(dest %in% top_dest$dest)
```

#### Choose all flights that go to the top ten destinations using semi_join

```{r top ten semi join}
flights %>% 
  semi_join(top_dest)
```

#### Choose all flights that do not go to the top ten destinations using anti_join

```{r not top ten}
flights %>%
  anti_join(top_dest)
```

### Practice

#### Exercise 1

Total number of flights from each airport with full name of airport

```{r total flights from each airport}
flights %>% 
  count(origin) %>% 
  left_join(airports, by = c("origin" = "faa")) %>% 
  select(
    Airport = name,
    Number_of_Flights = n)
```

#### Exercise 2

Destination airports with the top 10 largest average arrival delays

```{r largest arrival delays}
flights %>% 
  group_by(dest) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE) %>% round(2)
  ) %>% 
  ungroup() %>% 
  arrange(delay %>% desc()) %>% 
  slice_head(n = 10) %>%
  left_join(airports, by = c("dest" = "faa")) %>% 
  select(
    Destination = name,
    Delay = delay
  )
  
```

#### Exercise 3

Average departure delay by airline and departure airport

```{r average departure delay}
flights %>% 
  group_by(carrier, origin) %>% 
  summarise(
    average_delay = mean(dep_delay, na.rm = TRUE) %>% round(2)
  ) %>% 
  ungroup() %>% 
  left_join(airlines, by = "carrier") %>%
  select(
    carrier = name,
    origin,
    average_delay
  ) %>% 
  left_join(airports, by = c("origin" = "faa")) %>% 
  select(carrier,
         airport = name,
         average_delay) %>% 
  arrange(average_delay %>% desc())

```

#### Exercise 4

-   Remove flights with missing tail numbers
-   Top 10 planes with the most flights
-   Add full carrier name

```{r most flown aircraft}
(top10_planes <- flights %>% 
  filter(!is.na(tailnum)) %>% 
  count(tailnum, sort = TRUE) %>% 
  slice_head(n = 10)
)

top10_planes %>% 
  left_join(flights) %>% 
  distinct(tailnum, .keep_all = TRUE) %>% 
  left_join(airlines) %>% 
  select(tailnum, airline = name, number_of_flights = n)
```

#### Exercise 5

Choose all flights originating at JFK going to the top 10 destinations for JFK

```{r all flights from JFK to top 10 destinations}
top10_destinations <- flights %>% 
  filter(origin == "JFK") %>% 
  count(dest, sort = TRUE)

flights %>% 
  filter(origin == "JFK") %>% 
  semi_join(top10_destinations)
```
