---
title: "Relational Data"
author: "Ray Nelson"
date: "24 Feb 2019"
output:
  html_document:
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries
library(tidyverse)
library(nycflights13)
```

## Overview of tables to be manipulated  

### Airlines Table

```{r airlines table}
airlines
```  

### Airports Table

```{r airports table}
airports
```  
### Planes Table  

```{r planes table}
planes
```  
### Weather Table  

```{r weather table}
weather  
```  

## Keys: Primary and Foreign  

### Check to see if keys are unique for planes and weather tibbles

```{r check planes and weather}
planes %>% 
  count(tailnum) %>% 
  filter(n > 1)

weather %>% 
  count(year, month, day, hour, origin) %>% 
  filter(n > 1)
```  

### Sometimes we don't have an explicit primary key  

```{r check flights}  

flights %>% 
  count(year, month, day, flight) %>% 
  filter(n > 1)

flights %>% 
  count(year, month, day, tailnum) %>% 
  filter(n > 1)
```  

## Mutating joins  

```{r abbreviated flights}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)

flights2
```  

### Add the full airline name to the flights2 data.  

```{r full arline name}
flights2 %>%
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")
```  

### Understanding Joins  

#### Create the two tibbles x and y  

```{r join tibbles}
x <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2",
  3, "x3"
)

y <- tribble(
  ~key, ~val_y,
  1, "y1",
  2, "y2",
  4, "y3"
)
```  

#### inner join
```{r inner join}
x %>% 
  inner_join(y, by = "key")
```  

#### left join
```{r left join}
x %>% 
  left_join(y, by = "key")
```  

#### right join
```{r right join}
x %>% 
  right_join(y, by = "key")
```  

#### full join
```{r full join}
x %>% 
  full_join(y, by = "key")
```  

### Duplicate Keys

```{r duplicate}
x <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2",
  2, "x3",
  3, "x4"
)

y <- tribble(
  ~key, ~val_y,
  1, "y1",
  2, "y2",
  2, "y3",
  3, "y4"
)

x %>%
  left_join(y, by = "key")
```  

### Defining the key columns  

```{r default key}
flights2 %>% 
  left_join(weather)
```  

### Using the by  

```{r use by}
flights2 %>% 
  left_join(planes, by = "tailnum")
```  

### Using the by with two different tables with different key names  

```{r different key names}
flights2
airports

flights2 %>% 
  left_join(airports, c("dest" = "faa"))

flights2 %>% 
  left_join(airports, c("origin" = "faa"))
```  

### Filtering joins  

```{r filtering joins}
top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest  

flights %>% 
  filter(dest %in% top_dest$dest)

flights %>% 
  semi_join(top_dest)

flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(tailnum, sort = TRUE)
```  