---
title: "Introduction"
author: "Ray"
date: "12/30/2020"
output:
  html_document:
    code_folding: "hide"
editor_options: 
  markdown: 
    wrap: 80
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
```

Data Science for Managers

The intersection of statistics, computer science, and domain knowledge defines
data science. The process whereby data scientists explore, wrangle, program,
model, and communicate quantitative relationships is greatly facilitated by the
statistical computer language R.

Learning Outcomes

1.  Explore data through visualization and transformation

2.  Organize data into constructs that are useful for visualizing and modeling

3.  Master the philosophy, basics, and components of the R language

4.  Communicate clearly the insights gleaned from data exploration

```{r plot}
ggplot(data = mpg, mapping = aes(x = displ,
                                 y = hwy,
                                 color = drv)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(
    title = "The effect of engine size on fuel efficiency",
    subtitle = "Larger engines have lower miles per gallon.",
    y = "Highway Miles Per Gallon",
    x = "Engine Displacement",
    color = "Drive Train"
  )
```
