---
title: "Midterm I Winter 2019"
author: "Data Science for Managers"
date: "Due by class: 2/19/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(nycflights13)
```  

Submit an R Markdown for the following two problems. In order to get full credit, your R Markdown document must knit properly.  

## Problem 1

Please write code that completes the following:

*  Create a new variable for the departure hour from the scheduled departure time variable
*  Calculate the number of flights that leave during each hour and each airport
*  Draw a bar chart for each of the three New York airports
*  Color the bars based on their origin airport
*  It is not required that your graph replicate the graph labels. We will cover that in Chapter 28.

```{r problem 1}

```

## Problem 2

Please write code that completes the following:

*  Determine the number of destinations for each NYC airport
*  Create a bar chart that depicts the number of destinations
*  Color the bars based on the departure airport
*  Order the bars from smallest to largest number of destinations (This is called a Pareto Chart)
*  It is not required that your graph replicate the graph labels. We will cover that in Chapter 28.

```{r, echo = FALSE}
url <- "http://govfinance.byu.edu/govfinance/classes/dataScience/homework/graphs/HW11Graph.png"
```

![](`r url`)

```{r Problem 2}

```

