---
title: "Select and Mutate"
author: "MPA 634: Data Science for Managers"
date: "27 Jan 2020"
output:
  html_document:
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(nycflights13)
```

## Give multiple ways to select dep_time, dep_delay, arr_time, and arr_delay from the flights data frame.  

```{r column selection alternatives}
variables <- c("dep_time", "dep_delay", "arr_time", "arr_delay")
departure_delays <- flights %>%
  select(variables)
departures_delays <- flights %>%
  select(dep_time, dep_delay, arr_time, arr_delay)
departures_delays <- flights %>%
  select(starts_with("dep"), starts_with("arr"))
departures_delays <- flights %>%
  select(ends_with("time"), ends_with("delay"))
departures_delays <- flights %>% 
  select(4, 6:7, 9)
```

## Calculate the difference between dep_time and sched_dep_time and compare the result with dep_delay. How do we explain the difference? 

*  Use transmute to calcuate the new time difference
*  Compare dep_time with your calculation to see where they are different
*  Choose the first 20 observations

```{r departure comparison}
departure_comparison <- flights %>%
  transmute(
    dep_time,
    sched_dep_time,
    dep_delay,
    time_difference = dep_time - sched_dep_time,
    same = (dep_delay == time_difference)
  )
head(departure_comparison, 20)
```

## Replicate the calculation of dep_delay

*  Create a new data frame (departure_calculations) from flights that  
   +  Converts dep_time into minutes from midnight
   +  Converts sched_dep_time into minutes from midnight  
*  View dep_time, sched_dep_time, del_delay and their calculated equivalents

```{r conversion of times}
departure_calculations <- flights %>% 
  transmute(
    dep_time,
    sched_dep_time,
    dep_delay,
    dep_minutes = dep_time %/% 100 * 60 + dep_time %% 100,
    sched_dep_minutes = sched_dep_time %/% 100 * 60 + sched_dep_time %% 100,
    dep_delay_calculated = dep_minutes - sched_dep_minutes
  )
departure_calculations %>% 
  head(5)
```  

## Transform scatterplot of diamonds

### Draw a scatter plot of price on diamond weights

*  Use the diamonds data frame  
*  The price of diamonds is given by price  
*  The weight of the diamonds is given by carat  
*  Use transparency to help with the overplotting of so many observations  
*  Add a loess smooth without the confidence interval  
*  Add a simple regression line without the confidence interval  

```{r original scatterplot, message = FALSE}
diamonds %>%
  ggplot(mapping = aes(x = carat, y = price)) +
  geom_point(alpha = 1 / 100) +
  geom_smooth(color = "red", se = FALSE) +
  geom_smooth(method = "lm", color = "blue", se = FALSE)
```

### Scale price and carat by subtracting the mean and dividing by the standard deviation

*  Create a new data frame that includes scaled values for price and carat
*  Draw a scatterplot with transparency, loess, and simple regression as before

```{r scaling, message = FALSE}
diamonds %>%
  transmute(
    scale_price = (price - mean(price, na.rm = TRUE)) / sd(price, na.rm = TRUE),
    scale_carat = (carat - mean(carat, na.rm = TRUE)) / sd(carat, na.rm = TRUE)
  ) %>%
  ggplot(aes(x = scale_carat, y = scale_price)) +
  geom_point(alpha = 1 / 100) +
  geom_smooth(color = "red", se = FALSE) +
  geom_smooth(method = "lm", color = "blue", se = FALSE)
```

### Try a logarithmic transformation of price and carat

*  Create a new data frame that includes logarithmic transformed variables for price and carat
*  Draw a scatterplot with transparency, loess, and simple regression as before

```{r logarithmic transformation, message = FALSE}
diamonds %>%
  transmute(log_price = log(price),
            log_carat = log(carat)) %>%
  ggplot(mapping = aes(x = log_carat, y = log_price)) +
  geom_point(alpha = 1 / 100) +
  geom_smooth(color = "red", se = FALSE) +
  geom_smooth(method = "lm", color = "blue", se = FALSE)
```

### We often don't need to create new variables because we have functions that accomplish transformations. Two of these often used transformations are the functions scale and log. We can actually use these transformations within ggplot.

```{r scatterplots of transformed variables, message = FALSE}
ggplot(data = diamonds, mapping = aes(x = scale(carat), y = scale(price))) +
  geom_point(alpha = 1/100) +
  geom_smooth(color = "red", se = FALSE) +
  geom_smooth(method = "lm", color = "blue", se = FALSE)

ggplot(data = diamonds, mapping = aes(x = log(carat), y = log(price))) +
  geom_point(alpha = 1/100) +
  geom_smooth(color = "red", se = FALSE) +
  geom_smooth(method = "lm", color = "blue", se = FALSE)
```

## Rank the diamonds by weight and price

*  Create two new variables called size_rank and price_rank where 1 means the largest
*  View the largest diamonds by size
*  View the largest diamonds by price

```{r rank the diamonds and find the largest}
largest <- diamonds %>%
  mutate(size_rank = min_rank(desc(carat)),
         price_rank = min_rank(desc(price)))

# Largest by size
largest %>%
  arrange(size_rank) %>%
  select(carat, size_rank)
head(10)

# Largest by price
largest %>%
  arrange(price_rank) %>%
  select(price, price_rank)
head(10)
```  