---
title: "Homework #19"
author: "Geospatial"
date: "Due: 6 April 2021"
output:
   html_document:
     code_folding: show
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
library(leaflet)
```

## Concepts

1.  Explain how an object of sf class is related to a data frame.
2.  Use Google to compare and contrast choropleths with cartograms.

## Interpretation of code

### Code chunk 1

```{r First code chunk}
sf::st_as_sf(maps::map('state', plot = FALSE, fill = TRUE)) %>% 
  ggplot(aes(fill = ID)) +
  geom_sf(alpha = 0.5, show.legend = FALSE) +
  colorspace::scale_fill_discrete_qualitative(palette = "cold")
```

### Code chunk 2

```{r Second code chunk}
quakes %>% 
  slice_head(n = 20) %>% 
  leaflet() %>% 
  addTiles() %>% 
  addCircleMarkers(
    radius = 5,
    color = "red",
    stroke = FALSE, fillOpacity = 0.5
  )
```

## Write Code

Airbnb

-   Download an airbnb dataset from Kaggle and import the data into R.
-   Use leaflet to plot the locations of some of the properties.
-   Choose a map from <http://leaflet-extras.github.io/leaflet-providers/preview/index.html> .
-   Draw a second map that clusters the points like we did with the quakes data frame.
