---
title: "Graphics for Communication: Presidential"
author: "MPA 634: Data Science for Managers"
date: "4 December 2019"
output: 
  html_document:
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
# Load libraries
library(tidyverse)
```  

## Presidential Graph

### Add President Trump

```{r add TRUMP}
trump <- tibble(name = "Trump",
                start = )
```


Here is the graph that we are going to create step by step
```{r presidential}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, color = party)) +
  geom_point(size = 4) +
  geom_segment(aes(xend = end, yend = id), size = 3) +
  scale_x_date(breaks = presidential$start, date_labels = "'%y") +
  scale_y_continuous(breaks = seq(34, 44, 1), labels = presidential$name) +
  scale_colour_manual(values = c("Republican" = "red", "Democratic" = "blue")) +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "bottom"
  ) +
  labs(
    title = "The political party of the elected US president regularly changes.",
    x = "",
    y = "",
    color = "Political Party"
  )
```  

#### What does the presidential tibble look like?  

```{r glimpse presidential}
glimpse(presidential)
```

#### Presidential id number

Add the presidential id number to the presidential tibble. Eisenhower was the 34th president.  

```{r president id number}
presidential %>%
  mutate(id = 33 + row_number())
```

#### Basic graph

Draw the basic graph

```{r basic presidential graph}
(basic <- presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, color = party)) +
    geom_point(size = 4)
)
```

#### Add the line segment that shows duration of each administration by using geom_segment  

```{r add duration}
(add_duration <- basic +
    geom_segment(aes(xend = end, yend = id), size = 3)
)
```

#### Change the scale of the x axis so that it labels the beginning of each administration.  
```{r x-axis labels}
(x_axis <- add_duration +
    scale_x_date(breaks = presidential$start, date_labels = "'%y")
)
```

#### Change the scale of the y axis so that it corresponds to each president and use each President's last name as the label   

```{r presidential names}
(y_axis <- x_axis +
  scale_y_continuous(breaks = seq(34, 44, 1), labels = presidential$name)
)
```

#### Change the color of the bars to be red for Republican and blue for Democrat 

```{r blue for democrat and red for republican}
(blue_red <- y_axis +
  scale_colour_manual(values = c("Republican" = "red", "Democratic" = "blue"))
)
```

#### Alter the grid lines so that the graph isn't so complex and busy and put the legend at the bottom of the graph.  

```{r alter the grid lines}
(alter_grid <- blue_red +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "bottom"
  )
)
```

#### Add titles and label to the graph so that it is informative  

```{r titles and labels}
alter_grid +
  labs(
    title = "The political party of the elected US president regularly changes.",
    x = "",
    y = "",
    color = "Political Party"
  )
```