---
title: "Geospatial Visualization"
author: "MPA 634: Data Science for Managers"
date: "8 Apr 2020"
output:
  html_document:
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(leaflet)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(maps)
library(ggspatial)
library(tools)
library(lwgeom)
library(googleway)
library(ggrepel)

theme_set(theme_bw())
load(url("http://govfinance.byu.edu/govfinance/classes/dataScience/lectures/scripts/Geospatial.RData"))
```
# Leaflet Examples

## Coordinates for Tanner Building
```{r Tanner Building}
latitude <- 40.2504
longitude <- -111.6525

(TNRB <- leaflet() %>% 
    addTiles() %>% 
    addMarkers(lng = longitude,
               lat = latitude,
               popup = "Marriott School of Business")
)
```

## Function to create a map
```{r leaflet map function}
leaflet_map <- function(longitude, latitude, popup){
  require(leaflet)
  leaflet() %>% 
    addTiles() %>% 
    addMarkers(lng = longitude, lat = latitude, popup = popup)
}
```

## Map of the Nelsons Home
### Coordinates for Nelson Home
### Latitude: 40.268831
### longitude: -111.644243

```{r Mookis hideout}
(Nelsons <- leaflet_map(longitude = -111.644243,
            latitude = 40.268831,
            popup = "Mooki Hideout")
)
```

## Change the map
```{r alternative map tiles}
TNRB %>% 
  addProviderTiles(provider = providers$MtbMap,
                   options = providerTileOptions(opacity = 0.1))
TNRB %>% 
  addProviderTiles(provider = providers$Stamen.Toner)
TNRB %>% 
  addProviderTiles(provider = providers$CartoDB.Positron)
```

## Use map overlays to create layers
```{r stack map layers}
TNRB %>%
  addProviderTiles(provider = providers$MtbMap,
                   options = providerTileOptions(opacity = 0.5)) %>%
  addProviderTiles(provider = providers$Stamen.TonerLines,
                   options = providerTileOptions(opacity = 0.5)) %>%
  addProviderTiles(provider = providers$Stamen.TonerLabels)
```

## Activity for breakout groups

* Use Google maps to find the coordinates for the Salt Lake City Temple
* Use Leaflet and the CartoDB.Positron provider to draw the map
* Locate and label the temple

```{r Salt Lake City Temple}
latitude <- 40.77066
longitude <- -111.891987
map_label <- "Salt Lake City Temple"

(TNRB <- leaflet() %>% 
    addProviderTiles(provider = providers$CartoDB.Positron) %>% 
    addMarkers(lng = longitude,
               lat = latitude,
               popup = map_label)
)
```

# Florida Choropleth

## Finished product
```{r finished choropleth}
world %>%
  ggplot() +
  geom_sf(fill = "antiquewhite1") +
  geom_sf(data = counties, aes(fill = area)) +
  geom_sf(data = states, fill = NA) +
  geom_sf(
    data = sites,
    size = 4,
    shape = 23,
    fill = "darkred"
  ) +
  geom_sf(data = flcities) +
  geom_text_repel(
    data = flcities,
    aes(x = lng, y = lat, label = city),
    fontface = "bold",
    nudge_x = c(1, -1.5, 2, 2, -1),
    nudge_y = c(0.25, -0.25, 0.5, 0.5, -0.5)
  ) +
  geom_label(
    data = states,
    aes(X, Y, label = ID),
    size = 5,
    fontface = "bold",
    nudge_y = states$nudge_y
  ) +
  scale_fill_viridis_c(trans = "sqrt", alpha = .4) +
  annotation_scale(location = "bl", width_hint = 0.4) +
  annotation_north_arrow(
    location = "bl",
    which_north = "true",
    pad_x = unit(0.75, "in"),
    pad_y = unit(0.5, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  coord_sf(xlim = c(-88, -78),
           ylim = c(24.5, 33),
           expand = FALSE) +
  labs(
    title = "Observation Sites",
    subtitle = "(2 sites in Palm Beach County, Florida)",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme(
    panel.grid.major = element_line(
      color = gray(0.5),
      linetype = "dashed",
      size = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  )
```

## Get basic background map data

```{r basic world map data}
world <- ne_countries(scale = "medium", returnclass = "sf")
```

## Draw the basic map

```{r basic map}
world %>% 
  ggplot() +
  geom_sf()
```

## Zoom in on Florida

```{r zoom in}
world %>% 
  ggplot() +
  geom_sf() +
  coord_sf(xlim = c(-88, -78),
           ylim = c(24.5, 33),
           expand = FALSE)
```

## Now add points for two sites
```{r two site points}
(sites <- tibble(longitude = c(-80.144005, -80.109),
                     latitude = c(26.479005, 26.83)))
world %>% 
  ggplot() +
  geom_sf() +
  geom_point(data = sites, aes(x = longitude, y = latitude),
             size = 4,
             shape = 23,
             fill = "darkred") +
  coord_sf(xlim = c(-88, -78),
           ylim = c(24.5, 33),
           expand = FALSE)
```

## Let's use **sf** rather than **geom_points** to graph sites
```{r same graph using sf}
(sites <- st_as_sf(sites, coords = c("longitude", "latitude"),
                   crs = 4326, agr = "constant"))
world %>%
  ggplot() +
  geom_sf() +
  geom_sf(
    data = sites,
    size = 4,
    shape = 23,
    fill = "darkred"
  ) +
  coord_sf(xlim = c(-88,-78),
           ylim = c(24.5, 33),
           expand = FALSE)
```

## Add states
```{r Add states}
states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
head(states)

states <- cbind(states, st_coordinates(st_centroid(states)))
states$ID <- states$ID %>% as.character() %>% toTitleCase()
head(states)

world %>%
  ggplot() +
  geom_sf() +
  geom_sf(data = states, fill = NA) +
  geom_text(data = states, aes(X, Y, label = ID), size = 5) +
  coord_sf(xlim = c(-88,-78),
           ylim = c(24.5, 33),
           expand = FALSE)
```

## Move the state labels

```{r move the state labels}
states$nudge_y <- -1
states$nudge_y[states$ID == "Florida"] <- 0.5
states$nudge_y[states$ID == "South Carolina"] <- -1.5

world %>% 
  ggplot() +
  geom_sf() +
  geom_sf(data = states, fill = NA) +
  geom_label(data = states,
             aes(x = X, y = Y, label = ID),
             size = 5,
             fontface = "bold",
             nudge_y = states$nudge_y) +
  coord_sf(xlim = c(-88,-78),
           ylim = c(24.5, 33),
           expand = FALSE)
```

## Counties (polygon data)

```{r counties data and overlay}

counties <- st_as_sf(map("county", plot = FALSE, fill = TRUE))
counties <- subset(counties, grepl("florida", counties$ID))
counties$area <- as.numeric(st_area(counties))
head(counties)

world %>% 
  ggplot() +
  geom_sf() +
  geom_sf(data = counties, fill = NA, color = gray(0.5)) +
  coord_sf(xlim = c(-88,-78),
           ylim = c(24.5, 33),
           expand = FALSE)
```


## Fill based on their area
```{r fill counties}
world %>% 
  ggplot() +
  geom_sf() +
  geom_sf(data = counties, aes(fill = area)) +
  scale_fill_viridis_c(trans = "sqrt", alpha = 0.4) +
  coord_sf(xlim = c(-88, -78),
           ylim = c(24.5, 33),
           expand = FALSE)
```

## Florida Cities data preparation
```{r Florida cities}

(flcities <- tibble(state = rep("Florida", 5),
                       city = c("Miami",
                                "Tampa",
                                "Orlando",
                                "Jacksonville",
                                "Sarasota"),
                       lat = c(25.7616798, 
                               27.950575,
                               28.5383355,
                               30.3321838,
                               27.3364347),
                       lng = c(-80.1917902,
                               -82.4571776,
                               -81.3792365,
                               -81.655651,
                               -82.5306527))
)
```
## Get the data for the cities from google

```{r Florida cities with help from google}
key <- "AIzaSyBj7F5KSag_NY51j2OqvjeGD-z9kajTNTw" # Google maps key
flcities <- data.frame(
  state = rep("Florida", 5),
  city = c("Miami",
           "Tampa",
           "Orlando",
           "Jacksonville",
           "Sarasota")
)
coords <- apply(flcities, 1, function(x) {
  google_geocode(address = paste(x["city"], x["state"], sep = ", "),
                 key = key)
})
flcities <- cbind(flcities, do.call(rbind,
                                    lapply(coords, geocode_coordinates)))

(flcities <-
    st_as_sf(
      flcities,
      coords = c("lng", "lat"),
      remove = FALSE,
      crs = 4326,
      agr = "constant"
    ))
```

## Simple Cities
```{r simple cities}
ggplot(data = world) +
  geom_sf() +
  geom_sf(data = counties,
          fill = NA,
          color = gray(.5)) +
  geom_sf(data = flcities) +
  geom_text(
    data = flcities,
    aes(x = lng, y = lat, label = city),
    size = 3.9,
    col = "black",
    fontface = "bold"
  ) +
  coord_sf(xlim = c(-88,-78),
           ylim = c(24.5, 33),
           expand = FALSE)
```

## Cities using ggrepel
```{r label cities using ggrepel}
ggplot(data = world) +
  geom_sf() +
  geom_sf(data = counties,
          fill = NA,
          color = gray(.5)) +
  geom_sf(data = flcities) +
  geom_text_repel(
    data = flcities,
    aes(x = lng, y = lat, label = city),
    fontface = "bold",
    nudge_x = c(1,-1.5, 2, 2,-1),
    nudge_y = c(0.25,-0.25, 0.5, 0.5,-0.5)
  ) +
  coord_sf(xlim = c(-88,-78),
           ylim = c(24.5, 33),
           expand = FALSE)
```

#E Final graph with compass orientation and scale
### Orientation and scale are inserted using annotations.

```{r final graph with orientation and scale}
world %>%
  ggplot() +
  geom_sf(fill = "antiquewhite1") +
  geom_sf(data = counties, aes(fill = area)) +
  geom_sf(data = states, fill = NA) +
  geom_sf(
    data = sites,
    size = 4,
    shape = 23,
    fill = "darkred"
  ) +
  geom_sf(data = flcities) +
  geom_text_repel(
    data = flcities,
    aes(x = lng, y = lat, label = city),
    fontface = "bold",
    nudge_x = c(1, -1.5, 2, 2, -1),
    nudge_y = c(0.25, -0.25, 0.5, 0.5, -0.5)
  ) +
  geom_label(
    data = states,
    aes(X, Y, label = ID),
    size = 5,
    fontface = "bold",
    nudge_y = states$nudge_y
  ) +
  scale_fill_viridis_c(trans = "sqrt", alpha = .4) +
  annotation_scale(location = "bl", width_hint = 0.4) +
  annotation_north_arrow(
    location = "bl",
    which_north = "true",
    pad_x = unit(0.75, "in"),
    pad_y = unit(0.5, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  coord_sf(xlim = c(-88, -78),
           ylim = c(24.5, 33),
           expand = FALSE) +
  labs(
    title = "Observation Sites",
    subtitle = "(2 sites in Palm Beach County, Florida)",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme(
    panel.grid.major = element_line(
      color = gray(0.5),
      linetype = "dashed",
      size = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  )
```

