---
title: "Vectors"
author: "Ray Nelson"
date: "11 March 2019"
output: 
  html_document:
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries
library(tidyverse)
```

## How to store and access information in R. We have the following possible objects:  

*  vectors (1 dimensional: character, logical, and numeric)
*  lists
*  tibbles (which are a special case of lists)  

## Create vectors in which we place information 

### Individuals associated with each section

```{r individuals associated with each section}
individuals1 <- c(
  "Derrik Adams",
  "Devin Adams",
  "Mike Carpenter",
  "Filipe Dias Bezerra",
  "Kiersten Ernstrom",
  "Ruben Ferreyra",
  "Jonathan Fortuna",
  "Matt Hubbard",
  "Teresa Hwang",
  "Nate Lambert",
  "Ryan Lancaster",
  "Matt Lipps",
  "Luciano Peixoto",
  "Alex Rwankuba",
  "Shantel Sanders",
  "Hayden Smith",
  "Uros Stampe",
  "Bruno Torres",
  "Tiago Triumpho"
)

individuals2 <- c(
  "Heidy Comish",
  "Eveline De Medeiros Miranda",
  "Kyle Finneman",
  "Chris Holdaway",
  "Taylor King",
  "Trevor Lake",
  "Travis Lovell",
  "Hope Morrison",
  "Sarah Nielsen",
  "You Yeong Oh",
  "Chad Olesiak",
  "Jonathan Perkins",
  "Jameson Ranck",
  "Seth Randall",
  "Amanda Spencer",
  "Joshua Thornock",
  "Matt Widmer",
  "Nathan Zick"
)
```

### Programs in which individuals can participate  

```{r programs}
program_names <- c("MISM", "MBA", "MPA")
program_codes <- c(1:3)
```

### Programs that correspond to each individual  

```{r individuals in programs}
programs1 <- c(2, 2, 2,	2, 1, 1, 2, 2, 3, 3, 3, 2, 2, 2, 3, 3, 2, 2, 2)
programs2 <- c(3, 3, 1, 2, 1, 2, 2, 3, 3, 3, 2, 3, 2, 2, 2, 2, 2, 1)
```

### Types of the objections that we have created

```{r types of objects}
typeof(individuals1)
typeof(programs1)
typeof(program_names)
typeof(program_codes)
```

## Organize our information into tibbles and lists

### Programs as a tibble and a list

```{r programs as a tibbles and a list}
(programs <- tibble(program_names, program_codes))
(programs_list <- list(program_names = program_names, program_codes = program_codes))
```

### Individuals as tibbles and lists

```{r individuals as tibble, error = TRUE}
length(individuals1)
length(individuals2)
(classes_tibble) <- tibble(individuals1, individuals2)
(class_list <- list(individuals1 = individuals1, individuals2 = individuals2))
```

### Individuals as a long tibble

```{r long individual tibble}
(section1 <- rep("section1", length(individuals1)))
(section2 <- rep("section2", length(individuals2)))

section1 <- tibble(individuals1, programs1, section1)
colnames(section1) <- c("individual", "program", "section")
section1

section2 <- tibble(individuals2, programs2, section2)
colnames(section2) <- c("individual", "program", "section")
section2

(classes <- rbind(section2, section1))
```

##### Add program identifier to the tibble with inner join and cleanup the tibble
```{r}
classes %>% 
  inner_join(programs, by = c("program" = "program_codes")) %>% 
  select(-program) %>% 
  rename(program = program_names)
```

## Information about objects

```{r object information}
colnames(classes)
dim(classes)
nrow(classes)
ncol(classes)
str(classes)
```

## Accessing information in lists

### Objects for use in this exercise include:

*  classes (tibble that was already created)
*  programs (tibble that was already created)
*  sections (list that needs to be created from section1 and section2)
*  a list called rdn_teaching which is a list that includes the above three objects

### Create the sections list

```{r sections list}
sections <- list(section1 = section1, section2 = section2)
``` 

### Type of the objects that we are going to put in rdn_teaching

```{r type of objects for rdn}
typeof(classes)
is_tibble(classes)
typeof(sections)
is_tibble(sections)
typeof(programs)
is_tibble(programs)
```

### Create the list rdn_teaching

```{r rdn_teaching creation}
rdn_teaching <- list(classes = classes, sections = sections, programs = programs)
```

### Investigate rdn_teaching

```{r rdn_investigation}
typeof(rdn_teaching)
is_tibble(rdn_teaching)
```

### Structure of the list  

```{r structure of rdn_teaching}
str(rdn_teaching)
```
### Accessing information in the first level of the list

```{r accessing elements in the rdn_list}
rdn_teaching$classes
rdn_teaching[[1]]
```

### Accessing information in a list that is within a list

```{r list within a list}
rdn_teaching$programs$program_names
rdn_teaching[[3]][[1]]
```  